---
title: "Beef"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)
```

In this file we use 2017 Census of Agriculture data to understand the expenditure patterns and land use of beef and dairy producers. While the focus of our analysis is New York, due to the potential for small sample sizes, we gather data for New York and the Middle Atlantic for Feedlot operations, the rest are for New York only.

## Import data

```{r}
library(tidyverse)
library(srvyr)
library(openxlsx)
library(broom)

# Load data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa_17.RData")
coa_17 <- coa_17 %>% as_survey(weights = weight)

# Import metadata 
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/meta_coa.RData")

# Import my functions 
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/myfunctions.R")

# Begin excel output file 
wb <- createWorkbook()

```

## Define variables

Dairy operations are defined as operations with milk cows kept for production of milk, include dry milk cows and milk heifers that had calved (k805\>0). While some dairy operations in NY are diversified, most only produce milk. We keep all operations in the sample, even if they are diversified.

-   Diary scale: 50-999, 100-499, and 500 or more dairy cows

    -   From USDA ERS Report [Consolidation in U.S. Dairy Farming](https://www.ers.usda.gov/webdocs/publications/98901/err-274.pdf?v=1435.3)

-   All data are presented grouped by these three scales and organic/non-organic for operations located in New York

-   Also include GCFI, total sales, dairy sales, dairy sales as a percent of total sales, and other operation characteristics

Beef operations are defined by beef operation type (defined below). We run separately for each beef operation type by operations with beef sales only (no diversified operations are included) and for those with at least 50% beef sales. All data is provided for New York, except for Feedlots, which is for New York and Middle Atlantic.

-   Beef operation type:

    -   Cow-calf only
    -   Feedlot only
    -   Other cattle

-   Scale

    -   Cow-calf: 1-49, 50-99, 100 or more head

        -   From USDA ERS report [The Diverse Structure and Organization of U.S. Beef Cow-Calf Farms](https://www.ers.usda.gov/webdocs/publications/44530/7611_eib73.pdf?v=513.8)
    -   Feedlot: 1 to 99, 100 - 199 , 200 or more steers and heifers on feed
    -   Other cattle : Same scale as cow-calf but based on number of other cattle (k1206-k1207)

```{r}

# Define operation type 
df <- coa_17  %>% 
  mutate(
    cowcalf = case_when(
      k804>0 & k805==0 & k1207==0~ 1, 
      TRUE ~ 0), 
    feedlot = case_when(
      k1207>0 & k805==0 & k804==0 ~ 1, 
      TRUE ~ 0), 
    catt_oth = case_when(
      cattle_sales>0 & k805==0 & 
        cowcalf==0 & feedlot==0 ~ 1, 
      TRUE ~ 0))

# Define scale
df <- df %>% 
  mutate(
    dairy_scale = case_when(
      k805>=50 & k805<100 ~ "50-99 dairy cows", 
      k805>=100 & k805<500 ~ "100 to 499 dairy cows",
      k805>=500 ~ "500 or more dairy cows",
      TRUE ~ NA),
    cowcalf_scale = case_when(
      k804>0 & k804<50 ~ "1 to 49 head", 
      k804>=50 & k804<100 ~ "50 to 99 head", 
      k804>=100 ~ "100 head or more", 
      TRUE ~ NA), 
    feedlot_scale = case_when(
      k1207 >0 & k1207<100 ~ "1 to 99 head", 
      k1207>=100 & k1207<200 ~ "100 to 199 head", 
      k1207>=200 ~ "200 head or more", 
      TRUE ~ NA), 
    catt_oth_count = k1206-k1207 + k808 + k810, 
    catt_oth_scale = case_when(
      catt_oth_count>0 & catt_oth_count<50 ~ "1 to 49 head", 
      catt_oth_count>=50 & catt_oth_count<100 ~ "50 to 99 head", 
      catt_oth_count>=100 ~ "100 head or more", 
      TRUE ~ NA))

# Define cattle only and diversified 
df <- df %>% 
  mutate(
    cattle_only_farm = case_when(
      cattle_sales == tvp ~ 1, 
      TRUE ~ 0), 
    cattle_diversified_farm = case_when(
      farmtype==11 ~ 1, 
      TRUE ~ 0))

# Define additional variables 
df <- df %>% 
  mutate(
    cattle_sold_less_500 = k808, 
    cattle_sold_500_plus = k810,
    other_cattle = k1206, 
    total_cattle = k803, 
    pasture_pct_acres = pasture_acres/operated_acres)

# Define metadata 
varlist <- df$variables %>% 
  select(year, dairy_scale:cattle_diversified_farm,
         -catt_oth_count,
         cattle_sold_less_500:pasture_pct_acres) %>% 
  names()

cols_remove <- "year"
m <- meta_select_vars(df$variables) 

# Add variable description 
m <- m %>% 
  mutate(
    variable_description = case_when(
      variable_name == "cowcalf" ~ "A cow-calf only operation", 
      variable_name == "feedlot" ~ "A feedlot only operation", 
      variable_name == "catt_oth" ~ "A backgrounder/stocker cattle operation", 
      variable_name == "dairy_scale" ~ "A factor variable with three levels: 50-99 dairy cows, 100 to 499 dairy cowsm 500 or more dairy cows",
      variable_name == "cowcalf_scale" ~ "A factor variable with three levels:1 to 49 head, 50 to 99 head, 100 head or more", 
      variable_name == "feedlot_scale" ~ "1 to 99 head sold, 100 to 199 head sold, 200 head or more on feed",
      variable_name == "catt_oth_scale"  ~ "A factor variable with three levels:1-49 head, 50-99 head, 100 head or more", 
      variable_name == "cattle_only_farm" ~ "An operation with only sales from cattle and total cattle sales of at least $10,000", 
      variable_name == "cattle_diversified_farm" ~ "An operation with at least 50% of sales from cattle and total cattle sales of at least $10,000", 
      variable_name == "cattle_sold_less_500" ~ "Nuber of calves sold or moved from the operation weighing less than 500 pounds",
      variable_name == "cattle_sold_500_plus" ~ "Cattle sold or moved from operation, including calves, weighing 500 pounds or more", 
      variable_name == "other_cattle" ~ "Number of other cattle and calves, including heifers that have not calves, steers, calves, and bulls", 
      variable_name == "total_cattle" ~ "Total number of cattle and calves on hand", 
      variable_name == "pasture_pct_acres" ~ "Pasture acres as a percent of total acres operated"))

# Add variable definition 
m <- m %>% 
  mutate(
    variable_definition = case_when(
      variable_name == "cowcalf" ~ "If k804>0 & k805==0 & k1207==0, then 1, otherwise 0", 
      variable_name == "feedlot" ~ "If k1207>0 & k805==0 & k804==0, then 1, otherwise 0", 
      variable_name == "catt_oth" ~ "A cattle operation that is not cow-calf only or feedlot only", 
      variable_name == "dairy_scale" ~ "Based on k805, equal to NA if not a diary operation or less than the smallest scale", 
      variable_name == "cowcalf_scale" ~ "Based on k804, equal to NA if not a cow-calf operation",
      variable_name == "feedlot_scale" ~ "Based on k1207, equal to NA if not a feedlot operation", 
      variable_name == "catt_oth_scale" ~ "Based on number of other cattle defined as k1206-k1207", 
      variable_name == "cattle_only_farm" ~ "k804>0 & k1379>=10000 & k1379==tvp", 
      variable_name == "cattle_diversified_farm" ~ "farmtype == 11 & k1379>=10000"))

# Add source and url 
m <- m %>% 
  mutate(
    source = "U.S. Department of Agriculture, National Agricultural Statistics Service, 2017 Census of Agriculture", 
    url = "https://www.nass.usda.gov/Publications/AgCensus/2017/index.php")

```

## Summary statistics

### Dairy

```{r}

# Define variable list
varlist <- df$variables %>% 
  select(
    dairy_sales, dairy_sales_tvp, dairy_herd, cattle_herd, 
    cattle_sold_less_500, cattle_sold_500_plus,
    other_cattle, total_cattle, 
    tvp, gcfi, gcfi_0_75, gcfi_75_350, gcfi_350_1M, gcfi_over_1M,
    gov_payments, crop_insurance, roa, opm, infi, 
    operated_acres, fullowner, partowner, fulltenant,
    beginning, whiteonlynonhis, female, young, senior, 
    vet, proportion_off_farm,  
    fert:otherexp, totvarexp_alt, 
    fert_alt_var_pct:otherexp_alt_var_pct,
    value_added_farm, value_added_sales, va_tvp, 
    rotational_grazing:veal, 
    permanent_pasture_acres:pasture, 
    fertilizer_cropland_acres:defoliation,
    solar:wind_rights, trucks:hay_balers, 
    pasture_acres, permanent_pasture_acres, 
    pasture_pct_acres) %>% 
  names()

# Define groupby variables
df_dairy <- df %>% 
  filter(!is.na(dairy_scale) & state_name == "New York") %>%
  mutate(groupbyvar = dairy_scale, 
         groupbyvar2 = organic_fct)

sum <- mean_se_median2(df_dairy)

addWorksheet(wb, "dairy")
writeData(wb, sheet = "dairy", sum)

# Add metadata 
meta_beef <- sum %>% 
  mutate(year = "2017") %>%
  select(year, variable_name) %>%
  distinct() %>%
  left_join(meta_coa, by = c("year", "variable_name")) %>% 
  filter(!is.na(variable_description)) %>% 
  bind_rows(m)

```

### Cattle

All data is split for beef operations that are diversified and those that sell beef only. Additionally they are split by organic/not organic and local/not local.

All data is for New York, except for feedlots which are provided for New York the Middle Atlantic (NY, PA, and NJ). Feedlot only operations are not in NY.

```{r}
# Define variable list
varlist <- df$variables %>% 
  select(
    cattle_sales, cattle_sales_tvp, cattle_herd, 
    cattle_sold_less_500, cattle_sold_500_plus,
    other_cattle, total_cattle, 
    tvp, gcfi, gcfi_0_75, gcfi_75_350, gcfi_350_1M, gcfi_over_1M,
    gov_payments, crop_insurance, roa, opm, infi, 
    operated_acres, fullowner, partowner, fulltenant,
    beginning, whiteonlynonhis, female, young, senior, 
    vet, proportion_off_farm,  
    fert:otherexp, totvarexp_alt, 
    fert_alt_var_pct:otherexp_alt_var_pct,
    value_added_farm, value_added_sales, 
    #va_tvp, 
    rotational_grazing:veal, 
    permanent_pasture_acres:pasture, 
    fertilizer_cropland_acres:defoliation,
    solar:wind_rights, trucks:hay_balers, 
    pasture_acres, permanent_pasture_acres, 
    pasture_pct_acres) %>% 
  names()

# Define a new df for the different types of cattle operations 
optype <- df$variables %>% 
  select(cowcalf, feedlot, catt_oth) %>% 
  names()
  
scale <- df$variables %>% 
  select(cowcalf_scale, feedlot_scale, catt_oth_scale) %>% 
  names()

# Use a function to create a dataframe for each operation type
# Create a list of these dataframes and then loop through them to create results
# Organic - All
myfun <- function(df, optype, scale){
  df_beef <- df %>% 
    filter(cattle_sales >=1000 & 
             state_name == "New York" & 
             !!sym(optype) ==1) %>% 
    mutate(
      groupbyvar = !!sym(scale), 
      groupbyvar2 = organic_fct)
  
  mean_se_median2(df_beef)
}

map2(optype, scale, ~{
  sum <- myfun(df, .x, .y)
  sheet_name <- paste("o_", .x)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, sum)
})

# Local - All 
myfun <- function(df, optype, scale){
  df_beef <- df %>% 
    filter(cattle_sales >=1000 & 
             state_name == "New York" & 
             !!sym(optype) ==1) %>% 
    mutate(
      groupbyvar = !!sym(scale), 
      groupbyvar2 = local_fct)
  
  mean_se_median2(df_beef)
}

map2(optype, scale, ~{
  sum <- myfun(df, .x, .y)
  sheet_name <- paste("l_", .x)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, sum)
})

# Organic - diversified
myfun <- function(df, optype, scale){
  df_beef <- df %>% 
    filter(cattle_sales >=1000 & 
             state_name == "New York" & 
             cattle_diversified_farm == 1 &
             !!sym(optype) ==1) %>% 
    mutate(
      groupbyvar = !!sym(scale), 
      groupbyvar2 = organic_fct)
  
  mean_se_median2(df_beef)
}

map2(optype, scale, ~{
  sum <- myfun(df, .x, .y)
  sheet_name <- paste("do_", .x)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, sum)
})

# Organic - not diversified
myfun <- function(df, optype, scale){
  df_beef <- df %>% 
    filter(cattle_sales >=1000 & 
             state_name == "New York" & 
             cattle_only_farm ==1 &
             !!sym(optype) ==1) %>% 
    mutate(
      groupbyvar = !!sym(scale), 
      groupbyvar2 = organic_fct)
  
  mean_se_median2(df_beef)
}

map2(optype, scale, ~{
  sum <- myfun(df, .x, .y)
  sheet_name <- paste("ndo_", .x)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, sum)
})

# Local - diversified
myfun <- function(df, optype, scale){
  df_beef <- df %>% 
    filter(cattle_sales >=1000 & 
             state_name == "New York" & 
             cattle_diversified_farm == 1 &
             !!sym(optype) ==1) %>% 
    mutate(
      groupbyvar = !!sym(scale), 
      groupbyvar2 = local_fct)
  
  mean_se_median2(df_beef)
}

map2(optype, scale, ~{
  sum <- myfun(df, .x, .y)
  sheet_name <- paste("dl_", .x)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, sum)
})

# Local - not diversified
myfun <- function(df, optype, scale){
  df_beef <- df %>% 
    filter(cattle_sales >=1000 & 
             state_name == "New York" & 
             cattle_only_farm == 1 &
             !!sym(optype) ==1) %>% 
    mutate(
      groupbyvar = !!sym(scale), 
      groupbyvar2 = local_fct)
  
  mean_se_median2(df_beef)
}

map2(optype, scale, ~{
  sum <- myfun(df, .x, .y)
  sheet_name <- paste("ndl_", .x)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, sum)
})
save(wb, file = "data_processed/beefNY.RData")

# Feedlot - All Middle Atlantic
df_beef <- df %>% 
  filter(census_divisions == "Middle Atlantic" & 
           feedlot ==1 ) %>% 
  mutate(
    groupbyvar = feedlot_scale)
sum <- mean_se_median(df_beef)
addWorksheet(wb, "feedlot_MA")
writeData(wb, "feedlot_MA", sum)

# Feedlot - Diversified Middle Atlantic
df_beef <- df %>% 
  filter(census_divisions == "Middle Atlantic" & 
           feedlot ==1 & 
           cattle_diversified_farm == 1) %>% 
  mutate(
    groupbyvar = feedlot_scale)
sum <- mean_se_median(df_beef)
addWorksheet(wb, "d_feedlot_MA")
writeData(wb, "d_feedlot_MA", sum)

# Feedlot - Cattle only Middle Atlantic
df_beef <- df %>% 
  filter(census_divisions == "Middle Atlantic" & 
           feedlot ==1 & 
           cattle_only_farm ==1) %>% 
  mutate(
    groupbyvar = feedlot_scale)
sum <- mean_se_median(df_beef)
addWorksheet(wb, "nd_feedlot_MA")
writeData(wb, "nd_feedlot_MA", sum)

save(wb, file = "data_final/beef.RData")

```

## Metadata and info

```{r}

# Select variables of interest from meta_coa
meta_beef <- meta_coa %>% 
  filter(year == 2017 & 
           variable_name %in% all_of(varlist))

# Add newly created variables 
meta_beef <- bind_rows(meta_beef, m)

# Add worksheet 
addWorksheet(wb, "metadata")
writeData(wb, "metadata", meta_beef)

# Add info
info <- 
  tibble(
    sheet_name = 
      c("dairy", 
        "o_ cowcalf", "o_ feedlot", "o_ catt_oht",
        "l_ cowcalf", "l_ feedlot", "l_ catt_oht",
        "do_ cowcalf", "do_ feedlot", "do_ catt_oht", 
        "ndo_ cowcalf", "ndo_ feedlot", "ndo_ catt_oht",
        "dl_ cowcalf", "dl_ feedlot", "dl_ catt_oht", 
        "ndl_ cowcalf", "ndl_ feedlot", "ndl_ catt_oht",
        "feedlot_MA", "d_feedlot_MA", "nd_feedlot_MA", 
        "metadata", "info"), 
    description = 
      c("Mean/se/median for select variables for diary operations by scale and by organic/non organic operations", 
        
        "Mean/se/median for select variables for cow-calf only operations by scale and by organic/non organic operations", 
        "Mean/se/median for select variables for feedlot only operations by scale and by organic/non organic operations",
        "Mean/se/median for select variables for other cattle operations by scale and by organic/non organic operations",
        
        "Mean/se/median for select variables for cow-calf only operations by scale and by local/non-local operations", 
        "Mean/se/median for select variables for feedlot only operations by scale and by local/non-loca operations",
        "Mean/se/median for select variables for other cattle operations by scale and by local/non-loca operations",
        
        "Mean/se/median for select variables for diversified cow-calf only operations by scale and by organic/non-organic operations", 
        "Mean/se/median for select variables for diversified feedlot only operations by scale and by organic/non-organic operations",
        "Mean/se/median for select variables for diversified other cattle operations by scale and by organic/non-organic operations",
        
        "Mean/se/median for select variables for non-diversified cow-calf only operations by scale and by organic/non-organic operations", 
        "Mean/se/median for select variables for non-diversified feedlot only operations by scale and by organic/non-organic operations",
        "Mean/se/median for select variables for non-diversified other cattle operations by scale and by organic/non-organic operations",

          "Mean/se/median for select variables for diversified cow-calf only operations by scale and by local/non-local operations", 
        "Mean/se/median for select variables for diversified feedlot only operations by scale and by local/non-local operations",
        "Mean/se/median for select variables for diversified other cattle operations by scale and by local/non-local operations",
        
        "Mean/se/median for select variables for non-diversified cow-calf only operations by scale and by local/non-local operations", 
        "Mean/se/median for select variables for non-diversified feedlot only operations by scale and by local/non-local operations",
        "Mean/se/median for select variables for non-diversified other cattle operations by scale and by local/non-local operations",
        
         "Mean/se/median for select variables for feedlot only operations by scale",
        "Mean/se/median for select variables for feedlot only diversified operations by scale",
        "Mean/se/median for select variables for feedlot only non-diversified operations by scale",
        
    "Description of all variables used in the analysis", 
    "Description of data included in each sheet"), 
    sample = c(
      "Dairy operations in New York",
      "Cow-calf only operations in New York", 
      "Feedlot only operations in New York", 
      "Other cattle operations in New York", 
      "Cow-calf only operations in New York", 
      "Feedlot only operations in New York", 
      "Other cattle operations in New York", 
      "Cow-calf only operations in New York", 
      "Feedlot only operations in New York", 
      "Other cattle operations in New York", 
      "Cow-calf only operations in New York", 
      "Feedlot only operations in New York", 
      "Other cattle operations in New York", 
      "Cow-calf only operations in New York", 
      "Feedlot only operations in New York", 
      "Other cattle operations in New York", 
      "Cow-calf only operations in New York", 
      "Feedlot only operations in New York", 
      "Other cattle operations in New York", 
      "Feedlot only operations in the Middle Atlantic (NY,PA,NJ)", 
      "Feedlot only operations in the Middle Atlantic (NY,PA,NJ)", 
      "Feedlot only operations in the Middle Atlantic (NY,PA,NJ)",
      NA, NA))

# Add worksheet 
addWorksheet(wb, "info")
writeData(wb, "info", info)

```

## Save results

```{r}
saveWorkbook(wb, "data_final/beef.xlsx", overwrite = TRUE)
save(wb, file = "data_final/wb.RData")
```
