---
title: "Dairy consolidation"
format: html
editor: visual
---

Use 2007, 2012, 2017 track over time to see farms that lost milk cows, add organic vs. not and compare, pull as percent that didn't survive and the percent that survived and lost milk cows, what proportion of cows that were dropped for those that lost cows.

Using value-added sales vs. no value-added sales would also be an interesting comparison, but 2017 is the first year this question was asked.

## Import data

```{r}
library(tidyverse)
library(srvyr)
library(openxlsx)
library(broom)

# Import 2007 and 2017 data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa.RData")
coa <- coa %>% filter(year !=2012)

# Import survival data frame from 2007-2017
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/survival/coa_07_17.RData")

coa_07_17 <- coa_07_17 %>% 
  as_survey(weights = weight.x)

# Import metadata 
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/meta_coa.RData")

# Import my functions 
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/myfunctions.R")

# Begin excel output file 
wb <- createWorkbook()

```

## Define variables

Here we define a value-added catetgory based on available data in 2007, which includes organic, direct-to-consumer sales, and/or agritourism.

```{r}

# define dairy scale
coa_07_17 <- coa_07_17 %>% 
  mutate(
    dairy_scale.x = case_when(
      k805.x>0 & k805.x<50 ~ "1-49 diary cows",
      k805.x>=50 & k805.x<100 ~ "50-99 dairy cows", 
      k805.x>=100 & k805.x<500 ~ "100 to 499 dairy cows",
      k805.x>=500 ~ "500 or more dairy cows",
      TRUE ~ NA))

coa <- coa %>% 
  mutate(
    dairy_scale = case_when(
      k805>0 & k805<50 ~ "1-49 dairy cows", 
      k805>=50 & k805<100 ~ "50-99 dairy cows", 
      k805>=100 & k805<500 ~ "100 to 499 dairy cows",
      k805>=500 ~ "500 or more dairy cows",
      TRUE ~ NA))

# Define dummy for operations that lost herdsize
coa_07_17 <- coa_07_17 %>% 
  mutate(
    dairy_shrink = case_when(
      k805.x<=k805.y ~ 1, 
      TRUE ~ 0),
    dairy_shrink_fct = case_when(
      dairy_shrink ==1 ~ "Decreased herd size", 
      TRUE ~ "Herd size increased or remained the same"), 
    dairy_herd_change = k805.x - k805.y)

```

## Define sample
We include dairy operations in New York. 

```{r}

# define data frame to include dairy operations (in 2007 for survival) in New York 
coa_07_17 <- coa_07_17 %>% 
  filter(
    state_name.x == "New York" & 
      dairy_farm.x == 1 ) 

coa <- coa %>% 
  filter(
    state_name == "New York" & 
      dairy_farm == 1 )

# Add metadata 
meta_dairy <- tibble(
  year = "2007-2017",
  variable_name = c("dairy_shrink", "dairy_shrink_fct", 
                    "dairy_herd_change", "survive"), 
  variable_description = c(
    "A 0/1 varialbe for a dairy operation that decreased herd size from 2007 to 2017",
    "A factor variable with two levels: Decreased herd size and Herd size increased or remained the same from 2007 to 2017", 
    "The change in herd size from 2007 to 2017",
    "An operation that survived from 2007 to 2017"),
  variable_definition = c(
    NA, NA, 
    "A positive value means herd size increased and a negative value means herd size decreased from 2007 to 2017", 
    "A 1 indicates the operation survived and a 0 indicates they did not survive"), 
  source = "U.S. Department of Agriculture, National Agricultural Statistics Service, 2007 and 2017 Census of Agriculture", 
  url = "https://agcensus.library.cornell.edu/census_year/2007-census/
|https://www.nass.usda.gov/Publications/AgCensus/2017/index.php")

                      
```

## Summary stats

### Dairy operations by year

Here we provide summary stats for number of diary operations and total and average herd size in 2007 and 2017, by organic and no organic.

```{r}

# define continous variables 
varlist <- coa$variables %>% 
  select(
    dairy_sales, dairy_sales_tvp, dairy_herd, cattle_herd, 
    tvp, gcfi, gov_payments, roa, opm, infi, 
    proportion_off_farm,  
    fert:otherexp, totvarexp_alt, 
    fert_alt_var_pct:otherexp_alt_var_pct) %>% 
  names()

varlist_cont <- varlist

# Define groupby variables
coa <- coa %>% 
  mutate(
    groupbyvar = year,
    groupbyvar2 = organic_fct, 
    groupbyvar3 = dairy_scale)

# Get averages for each variable cross tab by organic for each year and just by year
sum <- mean_se_median(coa)
sum2 <- mean_se_median2(coa)
sum3 <- mean_se_median3(coa)

# Save worksheet
addWorksheet(wb, "cont_all")
writeData(wb, "cont_all", sum)

addWorksheet(wb, "cont_organic")
writeData(wb, "cont_organic", sum2)

addWorksheet(wb, "cont_organic_scale")
writeData(wb, "cont_organic_scale", sum3)

save(wb, file = "data_processed/dairy_iterim.RData")

# Define categorical variables 
varlist <- coa$variables %>% 
  select(
    differentiated_farm,
    gcfi_0_75, gcfi_75_350, gcfi_350_1M, gcfi_over_1M,
    gov_payments_cat, crop_insurance, 
    roa_crit_zone, opm_crit_zone, 
    operated_acres, fullowner, partowner, fulltenant,
    beginning, whiteonlynonhis, female, young, senior, 
    vet) %>% 
  names()

varlist_cat <- varlist

# Get totals for each variable cross tab by organic for each year and just by year
sum <- summary_total(coa)
sum2 <- summary_total2(coa)
sum3 <- summary_total3(coa)

# Save worksheet
addWorksheet(wb, "cat_all")
writeData(wb, "cat_all", sum)

addWorksheet(wb, "cat_organic")
writeData(wb, "cat_organic", sum2)

addWorksheet(wb, "cat_organic_scale")
writeData(wb, "cat_organic_scale", sum3)

save(wb, file = "data_processed/dairy_iterim.RData")

```

### Summary stats for those that did and did not survive

Include all of the same variables as before plus expansion/contraction for those that survived.

```{r}

# add a .x to all variables listed above
varlist_cont <- paste0(varlist_cont, ".x")
varlist_cat <- paste0(varlist_cat, ".x")
  
# Define groupby variables
varlist <- varlist_cont
coa_07_17 <- coa_07_17 %>% 
  mutate(
    groupbyvar = survive,
    groupbyvar2 = organic_fct.x, 
    groupbyvar3 = dairy_scale.x, 
    weight = weight.x)

# get summary stats 
sum <- mean_se_median(coa_07_17)
sum2 <- mean_se_median2(coa_07_17)
sum3 <- mean_se_median3(coa_07_17)

# Save worksheet
addWorksheet(wb, "survive_cont_all")
writeData(wb, "survive_cont_all", sum)

addWorksheet(wb, "survive_cont_organic")
writeData(wb, "survive_cont_organic", sum2)

addWorksheet(wb, "survive_cont_organic_scale")
writeData(wb, "survive_cont_organic_scale", sum3)

save(wb, file = "data_processed/dairy_iterim.RData")

# Totals for categorical variables 
varlist <- varlist_cat

sum <- summary_total(coa_07_17)
sum2 <- summary_total2(coa_07_17)
sum3 <- summary_total3(coa_07_17)

# Save worksheet
addWorksheet(wb, "survive_cat_all")
writeData(wb, "survive_cat_all", sum)

addWorksheet(wb, "survive_cat_organic")
writeData(wb, "survive_cat_organic", sum2)

addWorksheet(wb, "survive_cat_organic_scale")
writeData(wb, "survive_cat_organic_scale", sum3)

```

# Herd size change

```{r}
# Define data frame for operations that survived
df <- coa_07_17 %>% 
  filter(survive==1)

# Define groupby variables 
df <- df %>% 
  mutate(
    groupbyvar = dairy_shrink_fct, 
    groupbyvar2 = organic_fct.x)

# Get totals for each categorical variable cross tab by organic/shrink
varlist <- varlist_cat

sum <- summary_total(df)
sum2 <- summary_total2(df)

addWorksheet(wb, "herd_change_cat")
writeData(wb, "herd_change_cat", sum)

addWorksheet(wb, "herd_change_cat_org")
writeData(wb, "herd_change_cat_org", sum2)

# Get mean heard size change 
varlist <- "dairy_herd_change"
mean <- mean_se_median(df)
mean2 <- mean_se_median2(df)

addWorksheet(wb, "herd_change")
writeData(wb, "herd_change", mean)

addWorksheet(wb, "herd_change_organic")
writeData(wb, "herd_change_organic", mean2)

save(wb, file = "data_processed/dairy_iterim.RData")
```

### Survival rates

```{r}

## Total number of operations that survived and didn't by organic/not
# Define variables of interest 
varlist <- c("dairy_farm.x")

# Define groupby variables and change weight.x to weight
df <- coa_07_17 %>% 
  mutate(
    groupbyvar = organic_fct.x, 
    groupbyvar2 = survive,
    weight = weight.x)

# Get totals for each variable cross tab by organic/survive
sum <- summary_total2(df)

# Save worksheet
addWorksheet(wb, "survive")
writeData(wb, "survive", sum)

saveWorkbook(wb, "data_processed/dairy_iterim.xlsx")
```

## Save
Write metadata and info and save final data frame. 

```{r}

# Get meta data for existing variables
varlist <- coa$variables %>% 
  select(
    differentiated_farm,
    gcfi_0_75, gcfi_75_350, gcfi_350_1M, gcfi_over_1M,
    gov_payments_cat, crop_insurance, 
    roa_crit_zone, opm_crit_zone, 
    operated_acres, fullowner, partowner, fulltenant,
    beginning, whiteonlynonhis, female, young, senior, 
    vet) %>% 
  names()
  
m <- tibble(varlist) %>% 
  rename(variable_name = varlist)
  
varlist <- coa$variables %>% 
  select(
    dairy_sales, dairy_sales_tvp, dairy_herd, cattle_herd, 
    tvp, gcfi, gov_payments, roa, opm, infi, 
    proportion_off_farm,  
    fert:otherexp, totvarexp_alt, 
    fert_alt_var_pct:otherexp_alt_var_pct) %>% 
  names()

m2 <- tibble(varlist) %>% 
  rename(variable_name = varlist)

#bind with coa metadata
m <- bind_rows(m, m2) %>% 
  left_join(meta_coa)
rm(m2)

# Drop 2012
m <- m %>% 
  filter(year!=2012)

# Add new variables 
meta_dairy <- bind_rows(meta_dairy, m)

# Add worksheets
addWorksheet(wb, "meta")
writeData(wb, "meta", meta_dairy)

# Add info 
sheet_name <- getSheetNames("data_processed/dairy_iterim.xlsx")
sheet_name <- tibble(sheet_name)
description <- tibble(
  description = c(
  "mean/se/median for all continous variables by year",
  "mean/se/median for all continous variables by year, organic/non-organic",
  "mean/se/median for all continous variables by year, organic/non-organic and scale",
  "total for all categorical variables by year",
  "total for all categorical variables by year, organic/non-organic",
  "total for all categorical variables by year, organic/non-organic and scale",
  "mean/se/median for all continous variables by survive/didn't survive",
  "mean/se/median for all continous variables by survive/didn't survive, organic/non-organic",
  "mean/se/median for all continous variables by survive/didn't survive, organic/non-organic and scale",
  "total for all categorical variables by survive/didn't survive",
  "total for all categorical variables by survive/didn't survive, organic/non-organic",
  "total for all categorical variables by survive/didn't survive, organic/non-organic and scale",
  "total for all categorical variables by change in herd size for operations that survived",
  "total for all categorical variables by change in herd size and organic for operations that survived",
  "mean/se/median of dairy herd change for operations that changed herd size", 
  "mean/se/median of dairy herd change for operations by organic/non-organic that changed herd size",
  "survival rates comparing dairy and non-dairy operations", 
  "description of all variables used in the analysis"))

sample <- tibble(
  sample = c(
    "Dairy operations in New York, 2007 and 2017",
    "Dairy operations in New York, 2007 and 2017",
    "Dairy operations in New York, 2007 and 2017",
    "Dairy operations in New York, 2007 and 2017",
    "Dairy operations in New York, 2007 and 2017",
    "Dairy operations in New York, 2007 and 2017",
    "Dairy operations in New York that did and did not survive from 2007-2017",
    "Dairy operations in New York that did and did not survive from 2007-2017",
    "Dairy operations in New York that did and did not survive from 2007-2017",
    "Dairy operations in New York that did and did not survive from 2007-2017",
    "Dairy operations in New York that did and did not survive from 2007-2017",
    "Dairy operations in New York that did and did not survive from 2007-2017",
    "Dairy operations in New York that survived from 2007-2017",
    "Dairy operations in New York that survived from 2007-2017",
    "Dairy operations in New York that survived from 2007-2017",
    "Dairy and non-dairy operations in New York that did and did not survive from 2007-2017", NA, NA))


info <- bind_cols(sheet_name, description, sample)

# Add worksheets
addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save 
saveWorkbook(wb, "data_final/dairy_consolidation.xlsx")

```