---
title: "03.beans - additional data pull"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)
```

In this file we use 2022 Census of Agriculture data to create representative groups of dry bean producers in the Northeast and add information to our previous request from 02.beans.qmd.Sample sizes are too small for New York. 

We keep operations that had positive dry bean harvested acres (could also have soy) and operations that had positive soy acres but no dry beans. For the latter set of operations, we drop those that produce primarily for feed (primary commodity beef/dairy) as we assume they would not switch to dry bean production.

All operations are required to have at least $1,000 in sales and located in the Northeast. 

## Import data

```{r}
rm(list=ls())

library(tidyverse)
library(srvyr)
library(openxlsx)
library(broom)

# Import 2022 census of ag data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa_22.RData")

# Import my functions 
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/mean_se_function.R")
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/total_function.R")

# Create workbook
wb <- createWorkbook()

# Define Commodity - dry beans or soybeans only and drop all other operations
df <- coa_22$variables %>% 
  mutate(
    dry_beans = case_when(
      k4554>0 ~ 1, 
      TRUE ~ 0), 
    soybeans_only = case_when(
      k4088>0 & dry_beans==0 ~ 1, 
      TRUE ~ 0))
rm(coa_22)
```

### Final data
We keep operations that had positive dry bean harvested acres (could also have soy) and operations that had positive soy acres but no dry beans and were not primary commodity beef/dairy. All operations are required to have at least $1,000 in sales and located in the Northeast. 


```{r}
# Keep operations with positive value of sales for drybeans or soy only no primary commodity animal 
df <- df %>% 
  filter(dry_beans==1 | 
           (soybeans_only==1 & !primary_commodity %in% 
              c("Livestock", "Dairy"))
  )

# Drop operations with less than $1,000 in sales
df <- df %>% 
  filter(tvp>=1000)

# Define US
df_us <- df

# Define northeast
df <- df %>% 
  filter(census_regions == "Northeast")


```

## Define variables

### MWBE

MWBE is defined as an operation that falls into any of the following:

-   Ethnicity: label a farm as "socially disadvantaged" if at least one decision-maker on the farm checked yes to "Hispanic, Latino, or Spanish origin"
-   Race: label a farm as "socially disadvantaged" if at least one decision-maker on the farm checked yes to Black or African America, American Indian or Alaska Native, Asian, Native Hawaiian or Other Pacific Islander.
-   "Majority female" operations -where at least ½ of those involved in making management decisions on the farm were female

```{r}

# Define majority female based on decision making
df <- df %>% 
  mutate(
    op1_female_dm = case_when(
      op1_female==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_female_dm = case_when(
      op2_female==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_female_dm = case_when(
      op3_female==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_female_dm = case_when(
      op4_female==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_female_dm = op1_female_dm + 
      op2_female_dm + op3_female_dm + op4_female_dm, 
    majority_female_dm = case_when(
      total_female_dm >= (number_operators/2) ~ 1, 
      TRUE ~ 0))

# Define hispanic decision maker
df <- df %>% 
  mutate(
     op1_hispanic_dm = case_when(
      op1_hispanic==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_hispanic_dm = case_when(
      op2_hispanic==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_hispanic_dm = case_when(
      op3_hispanic==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_hispanic_dm = case_when(
      op4_hispanic==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_hispanic_dm = op1_hispanic_dm + 
      op2_hispanic_dm + op3_hispanic_dm + op4_hispanic_dm, 
    hispanic_dm = case_when(
      total_hispanic_dm>0 ~ 1, 
      TRUE ~ 0))

# Define black decision maker
df <- df %>% 
  mutate(
     op1_black_dm = case_when(
      op1_black==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_black_dm = case_when(
      op2_black==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_black_dm = case_when(
      op3_black==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_black_dm = case_when(
      op4_black==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_black_dm = op1_black_dm + 
      op2_black_dm + op3_black_dm + op4_black_dm, 
    black_dm = case_when(
      total_black_dm>0 ~ 1, 
      TRUE ~ 0))

# Define asian decision maker
df <- df %>% 
  mutate(
     op1_asian_dm = case_when(
      op1_asian==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_asian_dm = case_when(
      op2_asian==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_asian_dm = case_when(
      op3_asian==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_asian_dm = case_when(
      op4_asian==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_asian_dm = op1_asian_dm + 
      op2_asian_dm + op3_asian_dm + op4_asian_dm, 
    asian_dm = case_when(
      total_asian_dm>0 ~ 1, 
      TRUE ~ 0))

# Define indian or opi decision maker
df <- df %>% 
  mutate(
     op1_indian_opi_dm = case_when(
      op1_indian_opi==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_indian_opi_dm = case_when(
      op2_indian_opi==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_indian_opi_dm = case_when(
      op3_indian_opi==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_indian_opi_dm = case_when(
      op4_indian_opi==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_indian_opi_dm = op1_indian_opi_dm + 
      op2_indian_opi_dm + op3_indian_opi_dm + op4_indian_opi_dm, 
    indian_opi_dm = case_when(
      total_indian_opi_dm>0 ~ 1, 
      TRUE ~ 0))

## Define MWBE
df <- df %>% 
  mutate(
    mwbe_fct = case_when(
      majority_female_dm==1 |
        hispanic_dm==1 | 
        black_dm==1 | 
        asian_dm==1 |
        indian_opi_dm ==1 ~ "MWBE", 
      TRUE ~ "Not MWBE"), 
    mwbe = case_when(
      mwbe_fct == "MWBE" ~ 1, 
      TRUE ~ 0))

```

### Scale and others

```{r}

# Scale - under $1M and over
df <- df %>% 
  mutate(
    sales_class_agg = case_when(
      gcfi>=1000000 ~ "GCFI $1 million or more", 
      TRUE ~ "GCFI less than $1 million"))

# Define net farm income per acre operated
df <- df %>% 
  mutate(
    infi_acre = infi/operated_acres)

# Production cost per acre
df <- df %>%
  mutate(
    exp_acre = exp/operated_acres)

# Harvested acres and yields
df <- df %>% 
  mutate(
    dry_beans_harvested_acres = k554,
    dry_beans_cwt = k555,
    dry_beans_yield = case_when(
      dry_beans_harvested_acres==0 ~ 0, 
      TRUE ~ dry_beans_cwt/dry_beans_harvested_acres),
    soybeans_harvested_acres = k88, 
    soybeans_bushels = k89, 
    soybeans_yield = csoyxxyd)
  

# Dummy for primary commodity livestock or dairy operations
df <- df %>% 
  mutate(
    livestock_dairy = case_when(
      livestock==1 | dairy == 1 ~ 1, 
      TRUE ~ 0))

# CRP acres (also includes WRP, FWP, CREP)
df <- df %>% 
  mutate(
    crp_fct = case_when(
      k683>0 ~ 1, 
      TRUE ~ 0), 
    crp_acres = k683)

# Current inventory is acres harvested*yield - used to inform max_inventory
df <- df %>% 
  mutate(
    current_inventory_drybeans = dry_beans_harvested_acres*dry_beans_yield, 
    current_inventory_soybeans = soybeans_harvested_acres*soybeans_yield)


# Define tenure variable as owns more than 50 percent of operated land
df <- df %>% 
  mutate(
    owns_morethan50pct = case_when(
      owned_operated>.50 ~ 1, 
      TRUE ~ 0)
    )

# Define proportion of croplandacres devoted to cover crop, reduced till and no till (defined as a percent of harvested and harevested + failed)
df <- df %>% 
  mutate(
    cropland_total = cropland_harvested_acres + cropland_failed_acres,
    cover_crop_pct_cropland = cover_crop_acres/cropland_total,
    no_till_acres_pct_cropland = no_till_acres/cropland_total,
    reduced_till_acres_pct_cropland = reduced_till_acres/cropland_total, 
    intensive_till_acres_pct_cropland = intensive_till_acres/cropland_total,
    cover_crop_pct_harvested_cropland = 
      cover_crop_acres/cropland_harvested_acres,
    no_till_acres_pct_harvested_cropland = 
      no_till_acres/cropland_harvested_acres,
    reduced_till_acres_pct_harvested_cropland = 
      reduced_till_acres/cropland_harvested_acres, 
    intensive_till_acres_pct_harvested_cropland = 
      intensive_till_acres/cropland_harvested_acres
  )

# Define operations with all organic sales
df <- df %>% 
  mutate(
    all_organic_sales = case_when(
      organic_sales_tvp==1 ~ 1, 
      TRUE ~ 0)
  )


```

## Summary stats

### MWBE, scale

```{r}

# Make survey object
df <- df %>% 
  as_survey(weight = weight)

# Define varlist
varlist <- df$variables %>% 
  select(
    starts_with("current_inventory"),
    exp_acre,
    operated_acres,
    dry_beans_harvested_acres,
    dry_beans_cwt,
    dry_beans_yield,
    soybeans_harvested_acres,
    soybeans_bushels, 
    soybeans_yield,
    young, beginning, livestock_dairy,  
    organic_farm, 
    no_till, reduced_till, intensive_till,
    no_till_acres, reduced_till_acres, intensive_till_acres,
    no_till_acres_pct_cropland, reduced_till_acres_pct_cropland, 
    intensive_till_acres_pct_cropland,
    crp_fct, 
    owns_morethan50pct,
    differentiated_farm,
    infi, exp, tvp) %>% 
  names()


# Define groupby list
groupbylist <- df$variables %>% 
  select(sales_class_agg, mwbe_fct) %>% 
  names()

# Define croplist
croplist <- c("dry_beans", "soybeans_only")

# Mean/se
sum <- map(croplist, function(croplist){
  df %>% 
    filter(!!sym(croplist)==1) %>%
    mean_se_function(varlist, groupbylist) %>% 
    mutate(crop = croplist)
}) %>% 
  bind_rows() %>% 
  select(crop, everything())

## Total 
total <- map(croplist, function(croplist){
  df %>% 
    filter(!!sym(croplist)==1) %>%
    total_function(varlist, groupbylist) %>% 
    mutate(crop = croplist)
}) %>% 
  bind_rows() %>% 
  select(crop, everything())

# Bind rows
sum <- bind_rows(sum, total)

# Add Worksheet
addWorksheet(wb, "summary")
writeData(wb, "summary", sum)
```

### Land use by organic

```{r}

# Land use by organic
varlist <- df$variables %>% 
  select(
    cover_crop_pct_cropland,
    no_till_acres_pct_cropland, 
    reduced_till_acres_pct_cropland, 
    intensive_till_acres_pct_cropland, 
    cover_crop_pct_harvested_cropland,
    no_till_acres_pct_harvested_cropland, 
    reduced_till_acres_pct_harvested_cropland, 
    intensive_till_acres_pct_harvested_cropland,
    organic_sales_tvp, all_organic_sales) %>% 
  names()

# Define groupby list
groupbylist <- df$variables %>% 
  select(organic_fct) %>% 
  names()

# Mean/se
sum <- map(croplist, function(croplist){
  df %>% 
    filter(!!sym(croplist)==1) %>%
    mean_se_function(varlist, groupbylist) %>% 
    mutate(crop = croplist)
}) %>% 
  bind_rows() %>% 
  select(crop, everything())

# Add Worksheet
addWorksheet(wb, "summary_landuse_organic")
writeData(wb, "summary_landuse_organic", sum)

```

### Soybeans cost of production

Keep operations with soy only to compute cost of production per acre. Not run the second time.

```{r, eval = FALSE}

# Keep operations with soybean production only, not primary commodity livestock/dairy
soy_US <- df %>% 
  filter(k4088==tvp & !primary_commodity %in% 
              c("Livestock", "Dairy"))

soy <- df %>% 
  filter(k4088==tvp & !primary_commodity %in% 
              c("Livestock", "Dairy"))

# Summary stats
# Define varlist
varlist <- c("exp_acre")

# Define groupbylist
groupbylist <- c("sales_class_agg", "mwbe_fct", "organic_fct")

# Average for US
sum_us <- soy_US %>% 
  mean_se_function(varlist, groupbylist) %>% 
  mutate(sample = "US")

# Average for NY by organic only
groupbylist <- "organic_fct"

sum_ne <- soy %>% 
  mean_se_function(varlist, groupbylist) %>% 
  mutate(sample = "Northeast")

# Average for US by organic only
sum <- soy_US %>% 
  mean_se_function(varlist, groupbylist) %>% 
  mutate(sample = "US")


sum <- bind_rows(sum_us, sum, sum_ne)
rm(sum_us, sum_ne)


# Add worksheet
addWorksheet(wb, "soybeans_production_cost")
writeData(wb, "soybeans_production_cost", sum)

```

## OLS

```{r}

# Define model
results <- lm(
  infi ~ mwbe + sales_class_agg + current_inventory_drybeans + dry_beans_harvested_acres + young + organic_farm + partowner + exp_acre,
  data = df$variables, 
  weight = weight)

# Tidy
tidied <- tidy(results)
glanced <- glance(results)


addWorksheet(wb, "ols_tidy")
writeData(wb, "ols_tidy", tidied)

addWorksheet(wb, "ols_glance")
writeData(wb, "ols_glance", glanced)
```

## Save

```{r}

# Add info
info <- tibble(
  sheet_name = c("summary_organic", "summary", 
                 "ols_tidy", "ols_glance"), 
  description = c(
    "Summary stats for dry bean and soybean producers by scale, MWBE",
    "Summary stats for dry bean and soybean producers by scale, MWBE, organic",
    "OLS regression results", 
    "OLS model statistics including number of observations (nobs)"), 
  sample = c(
    "Dry bean and soybean producers in the Northeast"))
 
# Add worksheet
addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save 
saveWorkbook(wb, "../data_final/04.beans_final_results.xlsx", 
             overwrite = TRUE)

```
