---
title: "01.Beef"
format: html
editor: visual
---
In this file we use 2022 Census of Agriculture data to create representative groups of dairy/beef producers in New York State. We define operations based on those with primary commodity dairy. 

## Import data

```{r}

rm(list=ls())

library(tidyverse)
library(srvyr)
library(openxlsx)
library(broom)

# Import 2022 census of ag data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa_22.RData")

# Import my functions 
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/mean_se_function.R")
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/total_function.R")

# Create workbook
wb <- createWorkbook()

# Define Commodity - dairy
df <- coa_22$variables %>% 
  filter(dairy==1)

# Define NY only
df <- df %>% 
  filter(state_name=="New York")

# Drop operations with less than $1,000 in sales
df <- df %>% 
  filter(tvp>=1000)

```

## Define variables

### MWBE

MWBE is defined as an operation that falls into any of the following:

-   Ethnicity: label a farm as "socially disadvantaged" if at least one decision-maker on the farm checked yes to "Hispanic, Latino, or Spanish origin"
-   Race: label a farm as "socially disadvantaged" if at least one decision-maker on the farm checked yes to Black or African America, American Indian or Alaska Native, Asian, Native Hawaiian or Other Pacific Islander.
-   "Majority female" operations -where at least ½ of those involved in making management decisions on the farm were female

```{r}

# Define majority female based on decision making
df <- df %>% 
  mutate(
    op1_female_dm = case_when(
      op1_female==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_female_dm = case_when(
      op2_female==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_female_dm = case_when(
      op3_female==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_female_dm = case_when(
      op4_female==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_female_dm = op1_female_dm + 
      op2_female_dm + op3_female_dm + op4_female_dm, 
    majority_female_dm = case_when(
      total_female_dm >= (number_operators/2) ~ 1, 
      TRUE ~ 0))

# Define hispanic decision maker
df <- df %>% 
  mutate(
     op1_hispanic_dm = case_when(
      op1_hispanic==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_hispanic_dm = case_when(
      op2_hispanic==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_hispanic_dm = case_when(
      op3_hispanic==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_hispanic_dm = case_when(
      op4_hispanic==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_hispanic_dm = op1_hispanic_dm + 
      op2_hispanic_dm + op3_hispanic_dm + op4_hispanic_dm, 
    hispanic_dm = case_when(
      total_hispanic_dm>0 ~ 1, 
      TRUE ~ 0))

# Define black decision maker
df <- df %>% 
  mutate(
     op1_black_dm = case_when(
      op1_black==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_black_dm = case_when(
      op2_black==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_black_dm = case_when(
      op3_black==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_black_dm = case_when(
      op4_black==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_black_dm = op1_black_dm + 
      op2_black_dm + op3_black_dm + op4_black_dm, 
    black_dm = case_when(
      total_black_dm>0 ~ 1, 
      TRUE ~ 0))

# Define asian decision maker
df <- df %>% 
  mutate(
     op1_asian_dm = case_when(
      op1_asian==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_asian_dm = case_when(
      op2_asian==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_asian_dm = case_when(
      op3_asian==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_asian_dm = case_when(
      op4_asian==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_asian_dm = op1_asian_dm + 
      op2_asian_dm + op3_asian_dm + op4_asian_dm, 
    asian_dm = case_when(
      total_asian_dm>0 ~ 1, 
      TRUE ~ 0))

# Define indian or opi decision maker
df <- df %>% 
  mutate(
     op1_indian_opi_dm = case_when(
      op1_indian_opi==1 & k1642==1 ~ 1, 
      TRUE ~ 0), 
    op2_indian_opi_dm = case_when(
      op2_indian_opi==1 & k1643==1 ~ 1, 
      TRUE ~ 0), 
    op3_indian_opi_dm = case_when(
      op3_indian_opi==1 & k1644==1 ~ 1, 
      TRUE ~ 0), 
    op4_indian_opi_dm = case_when(
      op4_indian_opi==1 & k1645==1 ~ 1, 
      TRUE ~ 0), 
    total_indian_opi_dm = op1_indian_opi_dm + 
      op2_indian_opi_dm + op3_indian_opi_dm + op4_indian_opi_dm, 
    indian_opi_dm = case_when(
      total_indian_opi_dm>0 ~ 1, 
      TRUE ~ 0))

## Define MWBE
df <- df %>% 
  mutate(
    mwbe_fct = case_when(
      majority_female_dm==1 |
        hispanic_dm==1 | 
        black_dm==1 | 
        asian_dm==1 |
        indian_opi_dm ==1 ~ "MWBE", 
      TRUE ~ "Not MWBE"), 
    mwbe = case_when(
      mwbe_fct == "MWBE" ~ 1, 
      TRUE ~ 0))

```

### Scale and others

```{r}

# Scale - under $1M and over
df <- df %>% 
  mutate(
    sales_class_agg = case_when(
      gcfi>=1000000 ~ "GCFI $1 million or more", 
      TRUE ~ "GCFI less than $1 million"))

# Define net farm income per acre operated
df <- df %>% 
  mutate(
    infi_acre = infi/operated_acres)

# Production cost per acre and per cow
df <- df %>%
  mutate(
    exp_acre = exp/operated_acres, 
    exp_cow = case_when(
      dairy_herd==0 ~ 0, 
      TRUE ~ exp/dairy_herd)
  )


# Define dairy as a percent of total sales
df <- df %>% 
  mutate(
    dairy_tvp = dairy_sales/tvp
    )

# Land use practices
df <- df %>% 
  mutate(
    drained_tile_acres = k3450, 
    drained_ditches_acres = k3451, 
    conservation_easement_acres = k3452
  )


# Ohter cattle herd
df <- df %>% 
  mutate(
    other_cattle_herd = k1206, 
    cattle_herd_500plus = k810 
  )
```

## Summary stats

### Organic, MWBE, Scale

```{r}

# Define varlist
varlist <- df %>% 
  select(
    organic_farm, local_farm, 
    operated_acres, owned_acres, rented_in_acres, rented_out_acres, 
    partowner, fullowner, fulltenant,
    totvarexp, exp, infi, exp_acre, exp_cow,
    beginning,
    cropland_harvested_acres, 
    cropland_failed_acres, 
    cropland_summer_fallow_acres, 
    cropland_idle_acres, 
    permanent_pasture_acres, 
    woodland_pasture_acres,
    other_pasture_acres,
    drained_tile_acres,
    drained_ditches_acres,, 
    conservation_easement_acres,
    no_till, reduced_till, intensive_till,
    no_till_acres, reduced_till_acres, intensive_till_acres,
    cattle_herd, dairy_herd, other_cattle_herd, cattle_herd_500plus, 
    rotational_grazing, veal, 
    fertilizer_cropland_acres, fertilizer_pastureland_acres, 
    manure_acres, organic_acres, 
    insects_acres, weeds_acres, nematodes_acres, 
    diseases_acres, defoliation_acres, 
    methane_digester) %>% 
  names()


# Define groupby list
groupbylist <- df %>% 
  select(sales_class_agg, mwbe_fct, organic_fct) %>% 
  names()

# Mean/se
sum <- df %>% 
  as_survey(weight = weight) %>% 
  mean_se_function(varlist, groupbylist)


# Add Worksheet
addWorksheet(wb, "summary_organic")
writeData(wb, "summary_organic", sum)

```

### MWBE, scale

```{r}

# Not by organic

# Define groupby list
groupbylist <- df %>% 
  select(sales_class_agg, mwbe_fct) %>% 
  names()

# Mean/se
sum <- df %>% 
  as_survey(weight = weight) %>%
  mean_se_function(varlist, groupbylist)

# Add Worksheet
addWorksheet(wb, "summary")
writeData(wb, "summary", sum)
```

## OLS

```{r}

# Define model
results <- lm(
  infi ~ mwbe + sales_class_agg + dairy_herd + beginning + organic_farm + partowner,
  data = df, 
  weight = weight)

# Tidy
tidied <- tidy(results)
glanced <- glance(results)


addWorksheet(wb, "ols_tidy")
writeData(wb, "ols_tidy", tidied)

addWorksheet(wb, "ols_glance")
writeData(wb, "ols_glance", glanced)
```

## Save

```{r}

# Add info
info <- tibble(
  sheet_name = c("summary_organic", "summary", 
                 "ols_tidy", "ols_glance"), 
  description = c(
    "Summary stats for dairy producers by scale, MWBE",
    "Summary stats for dairy producers by scale, MWBE, organic",
    "OLS regression results", 
    "OLS model statistics including number of observations (nobs)"), 
  sample = c(
    "Dairy producers in New York"))
 
# Add worksheet
addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save 
saveWorkbook(wb, "../data_final/beef_results.xlsx", 
             overwrite = TRUE)

```
