### Description of Data Sources

All of the data from this file will be used to inform the supply chain of the ABM model.

**Wholesalers and Distributors**
We define the list of wholesale/distributor businesses from the SimplyAnalytics database of businesses located in New York State (NYS) that could carry beans. We assume that any wholesaler/distributor can carry MWBE identified product or local identified product. We also assume that wholesalers in NYS can supply all product demanded by New York City (NYC). 

The only businesses that are explicity included in the model are those that are identified as MWBE businesses or those that sold to NYC previously. 

**Aggregators/cleaners and processors**
We define the list of aggregator/cleaners and processors based on businesses that...

1.  were interviewed and/or identified by research team as key businesses in the bean supply chain (DryBean_SupplyChain_Data)
2.  sold to NYC previously and are located in NYS or located outside of NYS but sells NYS source-identified or NYS MW/BE product
    -   sells NYS source-identified product is defined in the NYC purchasing data by the variable "NY state spend" which indicates if an item is grown, processed, manufactured, or distributed by businesses located within New York State. The data is such that we cannot identify which business along the supply chain was a NYS product as the origin detail (processor), distributor, and vendor are all listed in one row where the "NY state spend" variable indicates yes/no all along the supply chain. We include all businesses along the supply chain that indicated at least one was a NYS product as NYS source identified.
3.  sold to NYC previously and are registered as a Minority/Women Owned Business (MWBE) through NYS and/or NYC
4.  were part of the list of the New York Food Products Database
    -   we include only those processors that produced legumes or plant-based protein

The following information is needed for all businesses:

-   Location (in NY or not)
-   Address
-   Role in the supply chain
-   Has the business sold to New York City
-   If the business sold to NYC, was the product from NY
-   MWBE or not
-   Works in the local and regional food supply chain

**Location** All data sets have information on if the business is located in NYS or not

**Address** We include an address for each business and then convert this to latitude and longitude coordinates to be included in the map. This data is not available for the majority of businesses so addresses are gathered and input manually.

**Role in the supply chain** For those businesses that are wholesalers/distributors, we have that information in all data sets. For processors, we do not have detailed enough information available in the SimplyAnalytics data and those are gathered by hand through web searches. For example with beans, we know which businesses process beans but we do not know if they are a primary or secondary processor. We do have that level of detail available in the data gathered through our interviews.

**Has the business sold to NYC** Any business that is in the NYC data we indicate has sold to NYC, we assume all other businesses have not.

**MWBE** These data are listed in the NYC data and the interview data. For the SimplyAnalytics data, in order to determine if businesses are MWBE, we use a [New York State database](https://ny.newnycontracts.com/FrontEnd/searchcertifieddirectory.asp) and a [NYC database](https://sbsconnect.nyc.gov/certification-directory-search/).

**Works with local farms** Businesses that sell locally are assumed to be those that were interviewed as part of this project and those found in the [Vendor Contact List in the NY Food Product Database](https://airtable.com/appNXQOhJImA3BVPc/shrCt48jrimPvahnI/tbliBTDzNm9JyA4vF/viwddIxh0H1NqtqHG?blocks=bipgl599cpG2pS1XM).

**Max inventory** This represents the maximum inventory that can flow through the businesses. If there is no supply contraints then we use the number of pounds of beans that NYC purchased as the max inventory with the goal that this number makes it so there is no constraint on how much the business can purchase in the model. 

**Organic**
We assume that all wholesalers/distributors can carry organic producers. Cleaners/aggregators and processors were identified by a web search if they can carry certified organic products. 

Note: For confidentiality reasons, we have removed business names, phone numbers, and addresses from the publicly available data. This code will not run as the source data is not available publicly, but the code can be used to understand our process.

### SimplyAnalytics
TODO: remove last 4 digits from the zipcode
This data is used for the wholesale busineses. 

[Simply Analytics](https://johnson.library.cornell.edu/database/simplyanalytics) is a web-based mapping application that allows users to map bushiness by name or by industry. Data are available to Cornell researchers through the Management Library. We collected data from (ADD YEAR) on the NAICS codes that are relevant to processing, distribution and wholesaling of dry beans, beef and leafy greens.

We use data aggregated at the 6-digit NAICS code.

The NAICS codes of interest for us include:

-   424420 Packaged Frozen Food Merchant Wholesalers
-   424410 General Line Grocery Merchant Wholesalers
-   424420 Packaged Frozen Food Merchant Wholesalers
-   424490 Other Grocery and Related Products Merchant Wholesalers
-   424510 Grain and Field Bean Merchant Wholesalers
-   484110 General Freight Trucking, Local
-   484121 General Freight Trucking, Long-Distance, Truckload
-   484122 General Freight Trucking, Long-Distance, Less Than Truckload
-   493110 General Warehousing and Storage
-   493120 Refrigerated Warehousing and Storage
-   493130 Farm Product Warehousing and Storage
-   493190 Other Warehousing and Storage

#### Import and clean data

```{r}
library(tidyverse)
library(openxlsx)
library(janitor)
library(readxl)
library(fuzzyjoin)

# Create workbook
wb <- createWorkbook()

# Define file path
file_path <- "../data_raw/ToddNAICS4.xlsx"

# Define sheet names of interest
sheet_names <- c("Wholesaling (42)", "Transportation (48)", 
                 "Warehousing & Storage (49)")

# Import data
naics_df <- sheet_names %>% 
  set_names() %>% 
  map_dfr(~suppressWarnings(read_excel(file_path, 
                      sheet = .x)) %>% 
            select(NAICS, Description, `Company Name`, `Street Address`, 
                   City, State, `Zip Code`) %>%
            mutate(
              `Zip Code` = as.character(`Zip Code`)), 
          .id = "sheet_name") %>% 
  clean_names()

# Filter NAICS of interest
wholesale_dist <- c("424420", "424410", "424420", "424490", 
                    "424510", "484110", "484121", "484122", 
                    "493110", "493120", "493130", "493190") 

naics_df <- naics_df %>% 
  filter(naics %in% wholesale_dist)


# Only keep businesses located in NY
naics_df <- naics_df %>% 
  filter(state=="NY")

```

#### Define variables

**Supply chain role** For the the processing sectors of the supply chain, we specify the commodity, and for the wholesale/distribution sectors we assume the sell all three commodities.

**Located in NY or not** We create a dummy variable with a 1 if in NY and 0 otherwise. All of these businesses are located in NYS. 

```{r}

# Supply chain role
naics_df <- naics_df %>% 
  mutate(
    wholesaler_distributor = 1)

# Define NY dummy (this will be used later, all businesses here are in NY)
naics_df <- naics_df %>% 
  mutate(
    ny = case_when(
      state=="NY" ~ 1, 
      TRUE ~ 0))

# Keep columns of interest
naics_df <- naics_df %>% 
  select(-c(naics, description, sheet_name))

# Add source
naics_df <- naics_df %>% 
  mutate(
    simply_analytics=1)

```

### New York City Purchasing
Overall, we use the NYC purchasing data to get a list of aggregator/cleaner and processor businesses that are located in NYS and have sold to NYC or are located outside of NYS but sell NYS source identified products to NYC. 

We also include a list of wholesalers that we will match with the Simply Analytics database to add additional information to the existing Simply Analytics data. 

Here we want the data from Citywide, which is what is downloaded when you download the full purchasing data from the [NYC Food Policy Dashboard](https://www.nyc.gov/site/foodpolicy/good-food-purchasing/citywidedata.page).

We keep all data relevant for the year 2023, based on primary food product category equal to legumes or meals that are meatless. In these data we have company names but we do not have information on their location.

The column called ny_state_spend_y_n indicates if an item is grown, processed, manufactured, or distributed by businesses located within New York State. The challenge is that the data includes up to three companies per line (origin_detail, distributor, and vendor) and for all three of these only one value for ny_state_spend_y_n. Therefore at least one, but not necessarily all of these businesses are in NYS. We want to identify businesses that sold NYS identified products to NYC, but using these data, we can't know which of the three businesses in each row was the one identified. As a first step, we assume all businesses in the row (i.e., origin_detail, distributor, and vendor) are assumed to supply NYS product.

We assume all businesses listed under origin_detail are processors and get the beans based on the products they sold and all other businesses are wholesaler/distributors. 

We want to include all businesses from this data set that are located in NYS and all out-of-state businesses that purchased NY identified products and sold to NYC (i.e., ny_state_spend_y_n=="y"), indicated by the variable "nys_product_purchase" equal to 1 if purchased source identified NY product and 0 otherwise.

While we do have a column in this dataset that indicates MWBE, we do not know from this data which business along the supply chain is MWBE. But since we have a list of all certified MWBE businesses for NYC and NYS, we rely on that list to confirm if a business is MWBE or not (rather than using the NYC purchasing database). Note that there are no bean processors that from the NYC data that were matched with the NYC/NYS MWBE certified data.  

We want to match data for all businesses listed. For each row there is a column called origin detail, distributor and vendor. We create a new column called company_name that includes all of the businesses listed in each column. We do not retain the columns pertaining to how much was purchased by NYC.

There are also a few businesses that sell meals with beans in them, specifically patties that we are including.

#### Import and clean
```{r}

# Import data
nyc_df <- read_xlsx("../data_raw/Public-Dashboard-Data-for-Download-FY19-23.xlsx", 
                 sheet = "Citywide") %>% 
  clean_names()

# filter data of interest
nyc_df <- nyc_df %>% 
  filter(time_period ==2023 & 
           primary_food_product_category == "Legumes" | 
           (primary_food_product_category=="Meals" & 
              str_detect(product_type, "MEATLESS"))
  )

# Max inventory for each business is what NYC purchased
nyc_bean_max <- nyc_df %>%
  summarise(sum(total_weight_in_lbs)) %>% 
  pull()

# Keep variables of interest
nyc_df <- nyc_df %>% 
  select(
    origin_detail, distributor, vendor,
    primary_food_product_category, ny_state_spend_y_n
    )

# Add supply chain role, ny, sold to nyc
nyc_df <- nyc_df %>% 
  mutate(nys_product_purchased = case_when(
           ny_state_spend_y_n=="Y" ~ 1, 
           TRUE ~ 0), 
         sold_to_NYC = 1) %>% 
  select(!c(ny_state_spend_y_n))
   
# Make all business names capitalized to join with Simply Analytics data
nyc_df <- nyc_df %>% 
  mutate(across(c(origin_detail, vendor, distributor),
                ~str_to_upper(.)))

# Get one column with all business names from origin_detail, vendor, distributor
nyc_df <- nyc_df %>% 
  pivot_longer(
    cols = !c(primary_food_product_category, nys_product_purchased, sold_to_NYC), 
    names_to = "name", 
    values_to = "company_name") 

# Add supply chain location 
nyc_df <- nyc_df %>% 
  mutate(
    beans_processor = case_when(
      name == "origin_detail" & primary_food_product_category %in% c("Legumes", "Meals") ~ 1, 
      TRUE ~ 0), 
    wholesaler_distributor = case_when(
      name %in% c("vendor", "distributor") ~ 1, 
      TRUE ~ 0))

# Some company names have location information included, drop everything after and including the first comma
nyc_df <- nyc_df %>% 
  mutate(company_name = str_remove(company_name, ",.*")) %>% 
  filter(company_name!="")

# Drop trailing spaces
nyc_df <- nyc_df %>% 
  mutate(
    company_name = str_trim(company_name, side = "right")
  )

# There are some companies where it looks like two companies (e.g., RIVER  / METRO COMMODITIES), for now these are included and can be changed when we do the manual data manipulation. 

# Keep distinct observations only and drop blanks
nyc_df <- nyc_df %>% 
  distinct(company_name, .keep_all = TRUE) %>% 
  filter(company_name != "(BLANK)")

# One company is listed multiple times but with slightly different names, keep only one
nyc_df <- nyc_df %>% 
  filter(!company_name %in% c("FURMANO", "FURMANO'S", "FURMANOS"))

# Drop columns we don't need
nyc_df <- nyc_df %>% 
  select(company_name, nys_product_purchased, sold_to_NYC, 
         beans_processor:last_col())

# Add data source
nyc_df <- nyc_df %>% 
  mutate(
    nyc_dashboard = 1)

```

#### Join NYC with SimplyAnalytics
Here we take the NYC wholesale data information and join it with the Simply Analytics data. 

##### NYC data is the base
We add addresses from the SA data to wholesalers that sold to NYC. And then we manually add in any missing addresses. This is one set of wholesale businesses that will be included in the model. 

We determine the max inventory to be a number where the quantitiy is unlimited. As a base, we add up the total pounds of products that NYC purchased. 

```{r}

# Keep NYC wholesale data only
nyc_wholesale_df <- nyc_df %>% 
  filter(wholesaler_distributor==1)

# Join with NYC data to get addresses from SimplyAnalytics data
nyc_wholesale_df <- stringdist_left_join(
  nyc_wholesale_df, naics_df, by = "company_name", 
  max_dist = 0.15, method = "jw")

# keep only one name and Drop duplicates 
nyc_wholesale_df <- nyc_wholesale_df %>% 
  select(-ends_with(".y"))

# remove .x at the end of some variables
nyc_wholesale_df <- nyc_wholesale_df %>% 
  rename_with(~str_remove(.x, "\\.x$"), ends_with(".x"))

# List missing
missing <- nyc_wholesale_df %>% 
  filter(is.na(street_address)) 

# Manually looked at missing, all but 1 located outside of NYS and none sold NYS source identified products, add missing addresses for relevant business
nyc_wholesale_df <- nyc_wholesale_df %>% 
  mutate(
    street_address = case_when(
      company_name == "TERI NICHOLS" ~ "10101c Avenue D",
      TRUE ~ street_address), 
    city = case_when(
      company_name == "TERI NICHOLS" ~ "Brooklyn",
      TRUE ~ city),
    state = case_when(
      company_name == "TERI NICHOLS" ~ "NY",
      TRUE ~ state),
    zip_code = case_when(
      company_name == "TERI NICHOLS" ~ "11236",
      TRUE ~ zip_code)
  )

# Drop businesses with missing information (i.e., out of NYS), all other businesses are in NYS
nyc_wholesale_df <- nyc_wholesale_df %>% 
  filter(!is.na(street_address))

# Keep columns of interest
nyc_wholesale_df <- nyc_wholesale_df %>% 
  select(company_name, street_address, 
         city, state, zip_code, 
         wholesaler_distributor, 
         nys_product_purchased, sold_to_NYC)


# add column for max inventory
nyc_wholesale_df <- nyc_wholesale_df %>% 
  mutate(
    max_inv_lbs_year = nyc_bean_max
  )

# Add NY variable
nyc_wholesale_df <- nyc_wholesale_df %>% 
  mutate(
    ny = 1)

# Save - this will be the list of wholesalers in the model
save(nyc_wholesale_df, file = "../data_processed/NYC_wholesalers.RData")

```

##### Simply Analystics is the base

```{r}

## To join with other data
# Keep NYC wholesale data only
wholesale_df <- nyc_df %>% 
  filter(wholesaler_distributor==1)

# Join NYC and SimplyAnalytics data using a fuzzy join, we want to keep all data and remove duplicates, there are no matches with the interview data
joined_df <- stringdist_left_join(
  naics_df, wholesale_df, by = "company_name", 
  max_dist = 0.15, method = "jw")

# Visually inspect, they are all the same, naics_df has address information for most businesses, so we want to keep the company name as listed in the Simply Analytics data
joined_df <- joined_df %>% 
  rename(company_name = company_name.x) 

# Keep Simply Analytics
joined_df <- joined_df %>% 
  select(-ends_with(".y")) 

# remove .x at the end of some variables
joined_df <- joined_df %>% 
  rename_with(~str_remove(.x, "\\.x$"), ends_with(".x"))

# rename
naics_df <- joined_df

rm(joined_df)

```

### Key supply chain businesses 
We import data on the businesses that were interviewed were interviewed and/or identified by research team as key businesses in the bean supply chain. All businesses are designated as local and regional food system businesses indicating they could distribute source identified product (lrfs_business==1).

These data include some businesses that were not interviewed. I only include the rows, Wild Kale through Vermont Bean Crafters (i.e., those with data included in a additional columns).

There are a few files with the same data, we use the file DryBean_SupplyChain_data.xlsx as it has the data that we need in the easiest format to use.

While there are important wholesalers and distributors listed here that can distribute NYS beans, because we assume that this is possible for all distributors, these are not explicitly modeled in our system. We mostly utilize this data for aggregator/cleaners and processors. 

#### Import and clean data
```{r}

# Import data collected by the qualitative team
qual_df <- read_xlsx("../data_raw/DryBean_SupplyChain_Data.xlsx", 
                     sheet = "Sheet1") %>% 
  clean_names() 


# Keep only data from those that were interviewed
qual_df <- qual_df %>% 
  filter(!is.na(agents_tolerence_for_risk))

# Change columns to be consistent (make business name capitalized to match with other data)
qual_df <- qual_df %>% 
  mutate(
    company_name = str_to_upper(business_name), 
    ny = case_when(
      in_state_out_of_state == "In-state" ~ 1, 
      TRUE ~ 0), 
    aggregator_cleaner = case_when(
      str_detect(supply_chain_segment, "Aggregator|Cleaner|cleaner") ~ 1, 
      TRUE ~ 0), 
    primary_processor = case_when(
      str_detect(supply_chain_segment, "Primary Processor") ~ 1, 
      TRUE ~ 0), 
    secondary_processor = case_when(
      str_detect(supply_chain_segment, "Secondary Processor") ~ 1, 
      TRUE ~ 0), 
    wholesaler_distributor = case_when(
      str_detect(supply_chain_segment, "Wholesaler|Distributor|copacker") ~ 1, 
      TRUE ~ 0))

# Keep data of interest
qual_df <- qual_df %>% 
  select(company_name, ny, aggregator_cleaner:wholesaler_distributor, 
         mwbe)


# Add local and regional food system business indicator for all businesses
qual_df <- qual_df %>% 
  mutate(
    lrfs_business=1)

# Add source
qual_df <- qual_df %>% 
  mutate(
    interview_data = 1)
```

#### Join to NYC/SA data

```{r}

# Do any of these businesses currently sell to NYC?
joined_df <- stringdist_inner_join(qual_df, nyc_df, 
                            by = "company_name", 
                            max_dist = 0.15, 
                            method = "jw")

# Only one company joined but has a different name. Easier to manually add
qual_df <- qual_df %>% 
  mutate(
    sold_to_NYC = case_when(
      company_name == "FURMANO FOODS" ~ 1, 
      TRUE ~ 0)
    )

rm(joined_df)

## Join with SimplyAnalytics data
joined_df <- stringdist_left_join(naics_df, qual_df,
                            by = "company_name", 
                            max_dist = 0.15, 
                            method = "jw")

# Keep data from SimplyAnalytics
joined_df <- joined_df %>% 
  rename(company_name = company_name.x) 

# Keep Simply Analytics columns when duplicated
joined_df <- joined_df %>% 
  select(-ends_with(".y")) 

# remove .x at the end of some variables
joined_df <- joined_df %>% 
  rename_with(~str_remove(.x, "\\.x$"), ends_with(".x"))

# rename
naics_df <- joined_df

rm(joined_df)

```

### NY Food Product Database

We use data the [NY Food Product Database](https://airtable.com/appNXQOhJImA3BVPc/shrCt48jrimPvahnI/tbliBTDzNm9JyA4vF/viwddIxh0H1NqtqHG?blocks=bipgl599cpG2pS1XM) to identify businesses that sell New York branded products (lrfs==1).

Bean processors include any manufacturer that lists the following values for Sub Category: "Legumes,Plant-Based Protein". There are no bean processors in this data base. However, we add one manually, Community Food Works (Milestone Mill) is a processor and distributor to include that was added after our data was available. 
We do not rely on the MWBE certification from this database as it is unclear whether is it for the processor or wholesaler. Rather we use the list of businesses certified MWBE by NYC and NYC. 

#### Import and clean data
```{r}

# Import data
ny_foods_df <- read_csv("../data_raw/NY Food Products - Approved-Grid View.csv") %>% 
  clean_names()

# Select data of interest
ny_foods_df <- ny_foods_df %>% 
  select(sub_category, manufacturer, distributors)

# Multiple distributors are listed in each row separated by a comma, make into a separate list
ny_foods_df <- ny_foods_df %>%
  separate_rows(distributors, sep = ",\\s*(?!\\s*Inc)") %>%
  mutate(distributors = str_trim(distributors), 
         distributors = str_remove_all(distributors, '["\']'))

# Define businesses that process beans
ny_foods_df <- ny_foods_df %>% 
  mutate(
    beans_processor = case_when(
      sub_category %in% c("Legumes,Plant-Based Protein") ~ 1, 
      TRUE ~ 0)) %>% 
  select(!sub_category)

# Add in Community Food Works (was added after our data)
ny_foods_df <- ny_foods_df %>% 
  add_row(
    manufacturer = "Community Food Works (Milestone Mill)",
    distributors = "Community Food Works (Milestone Mill)",
    beans_processor = 1
  )

# Keep only wholesale/distributors
distributors <- ny_foods_df  %>%
  select(distributors) %>%
  mutate(wholesaler_distributor = 1) %>% 
  distinct(.keep_all = TRUE) %>% 
  rename(company_name = distributors)

# Add processors
processors <- ny_foods_df  %>%
  filter(beans_processor==1) %>%
  select(manufacturer, beans_processor) %>%
  distinct(.keep_all = TRUE) %>% 
  rename(company_name = manufacturer)

# combine 
ny_foods_df <- full_join(distributors, processors) %>% 
  mutate(across(c(beans_processor, wholesaler_distributor), 
                ~case_when(is.na(.) ~ 0, 
                           TRUE ~ .))
  )

rm(distributors, processors)

# Make upper case and add lrfs column
ny_foods_df <- ny_foods_df %>% 
  mutate(company_name = toupper(company_name), 
         lrfs_business = 1)

# Filter if missing
ny_foods_df <- ny_foods_df %>% 
  filter(!is.na(company_name))

# Add source
ny_foods_df <- ny_foods_df %>% 
  mutate(
    food_product_database = 1
  )

```

#### Join to SimplyAnalytics
Here we take the NY food product data base wholesale data information and join it with the Simply Analytics data. 

```{r}

# Keep  wholesale data only
wholesale_df <- ny_foods_df %>% 
  filter(wholesaler_distributor==1)

# Join with SimplyAnalytics data using a fuzzy join
joined_df <- stringdist_left_join(
  naics_df, wholesale_df, by = "company_name", 
  max_dist = 0.15, method = "jw")

# Visually inspect, they are all the same, naics_df has address information for most businesses, so we want to keep the company name as listed in the Simply Analytics data
joined_df <- joined_df %>% 
  rename(company_name = company_name.x) 

# Keep lrfs from NY database
joined_df <- joined_df %>% 
  select(-lrfs_business.x) %>% 
  rename(lrfs_business = lrfs_business.y)

# Keep Simply Analytics
joined_df <- joined_df %>% 
  select(-ends_with(".y")) 

# remove .x at the end of some variables
joined_df <- joined_df %>% 
  rename_with(~str_remove(.x, "\\.x$"), ends_with(".x"))

# remove columns we don't need
joined_df <- joined_df %>%
  select(!c(beans_processor, primary_processor,
            secondary_processor, aggregator_cleaner))

# rename
naics_df <- joined_df

rm(joined_df)

```

### MWBE
We join the NYS and NYC MWBE certified businesses with all of our data sets. 

New York State provides a database with all [New York State Contractors that are certified MWBE](https://ny.newnycontracts.com/FrontEnd/searchcertifieddirectory.asp) and by [New York City](https://sbsconnect.nyc.gov/certification-directory-search/). We download the entire directory from both. 

We select the following [commodity codes](https://www.nyc.gov/assets/sbs/downloads/pdf/businesses/certification/new-york-city-commodity-code-listing.pdf):

-   019 - AGRICULTURAL CROPS AND GRAINS INCLUDING FRUITS, MELONS, NUTS
-   375 - FOODS: BAKERY PRODUCTS (FRESH)
-   380 - FOODS: DAIRY PRODUCTS (FRESH)
-   385 - FOODS, FROZEN
-   390 - FOODS: PERISHABLE
-   393 - FOODS: STAPLE GROCERY AND GROCER'S MISCELLANEOUS ITEMS

#### Import data
```{r}

# Import MWBE directories, keep data of interest and make capital
mwbe_ny <- read_csv("../data_raw/Directory_2025-05-30_862.csv", 
                    skip = 5) %>% 
  clean_names() %>%
  select(company_name) %>% 
  mutate(company_name = str_to_upper(company_name))

mwbe_nyc <- read_xlsx("../data_raw/OnlineDirectory06042025.xlsx", 
                      skip = 6) %>% 
  clean_names() %>% 
  mutate(company_name = str_to_upper(vendor_formal_name)) %>% 
  select(company_name)

# Join data (convert to UTF-8 format in order to join)
mwbe <- full_join(mwbe_ny, mwbe_nyc) %>% 
  mutate(mwbe = 1, 
         company_name = iconv(company_name, from = "", to = "UTF-8"))

rm(mwbe_ny, mwbe_nyc)

```

#### Join to SimplyAnalytics 

```{r}

# Join with other data so see if overlap
joined_df <- stringdist_inner_join(naics_df, mwbe, 
                                       by = "company_name",
                                       max_dist = 0.03, 
                                       method = "jw")

joined_df <- joined_df %>% 
  select(company_name.x, company_name.y, everything())

joined_df <- joined_df %>% 
  rename(company_name = company_name.x) %>% 
  distinct(company_name)
      
# Create a list of mwbe businesses and use that to add mwbe to df
mwbe_list <- joined_df %>% 
  pull(company_name)

naics_df <- naics_df %>% 
  mutate(
    mwbe = case_when(
      company_name %in% mwbe_list ~ 1, 
      TRUE ~ 0))

# Keep MWBE only
naics_df <- naics_df %>%
  filter(mwbe==1) %>%
  arrange(company_name)

# Drop duplicates, unless they have separate addresses
naics_df <- naics_df %>% 
  distinct(.keep_all = TRUE)


# Fill NAs with 0
naics_df <- naics_df %>% 
  mutate(across(nys_product_purchased:last_col(), 
                ~case_when(is.na(.) ~ 0, 
                           TRUE ~ .)))

# Make zip code 5 digits
naics_df <- naics_df %>% 
  mutate(zip_code = str_sub(zip_code, 1,5))

# Keep only columns of interest
naics_df <- naics_df %>% 
  select(
    company_name:ny, 
    nys_product_purchased, sold_to_NYC, 
    mwbe, lrfs_business)

# Save  - this will be wholesalers in the database
save(naics_df, file = "../data_processed/MWBE_wholesalers.RData")

```
#### Join to NYC processors
Here we match the data to the processor data only and find no matches. 

All NYC purchasing data from a wholesaler will be in the SimplyAnalytics data. 

```{r}

# No matches
joined_df <-  stringdist_inner_join(
  ny_foods_df %>%
    filter(beans_processor==1), 
  mwbe, by = "company_name",
  max_dist = 0.2, method = "jw")

```

#### Join to NY Food Product Dashboard data
Here we match the data to the processor data only and find no matches. 

All wholesaler will be in the SimplyAnalytics data. 

```{r}

joined_df <-  stringdist_inner_join(
  ny_foods_df %>%
    filter(beans_processor==1), 
  mwbe, by = "company_name",
  max_dist = 0.2, method = "jw")

```

### Interim Final data
Here we keep wholesaler data from SimplyAnalytics (with any additional information from other data sets included) and processor data from all other data sources. 

Here we add in the columns for the address data that will be added manually. At this step, the data is saved as an excel file and then addresses and other missing information are manually added by the qualitative team. After these additions, the data is imported, geocoded and then exported in its final format. 

```{r}

# Drop wholesale data from the NYC data
final_nyc_df <- nyc_df %>% 
  filter(beans_processor==1)

# Drop wholesale data from the NY food product data
final_ny_foods_df <- ny_foods_df %>% 
  filter(beans_processor==1)

# Drop wholesale data from the interview data
final_qual_df <- qual_df %>% 
  filter(aggregator_cleaner==1 | 
           primary_processor==1 | 
           secondary_processor==1)

# Combine processor data from all sources
df <- bind_rows(final_nyc_df, 
                final_ny_foods_df, 
                final_qual_df)

rm(final_ny_foods_df, final_ny_foods_df, final_qual_df)

# Arrange
df <- df %>% 
  arrange(company_name)

#  Manually drop duplicates with different names
df <- df %>% 
  filter(!company_name %in% c("DR PRAEGERS", 
                              "GARDEIN HAGERSTOWN, BRITISH COLOMBIA, CANADA MD", 
                              "HOUSE FOODS AMERICA", 
                              "PORT ROYAL SALES", 
                              "SYS CLS", 
                              "USDA")) %>% 
  filter(!str_detect(company_name, "BRITISH COLOMBIA")) %>%
  mutate(
    company_name = case_when(
      company_name == "WILD KALE" ~ "WILDKALE", 
      TRUE ~ company_name)
  ) 

# Manually fix duplicates
df <- df %>% 
  filter(company_name!="FURMANO FOODS") %>% 
  add_row(
    company_name="FURMANO FOODS", 
    nys_product_purchased=1, 
    sold_to_NYC=1, 
    beans_processor=1, 
    wholesaler_distributor=0, 
    nyc_dashboard=1, 
    lrfs_business=1, 
    food_product_database=NA, 
    ny = 0, 
    aggregator_cleaner = 0,
    primary_processor = 1,
    secondary_processor=0,
    mwbe = 0, 
    interview_data=1
  )

df <- df %>% 
  filter(company_name != "GENESEE VALLEY BEAN CO") %>%
  add_row(
    company_name = "GENESEE VALLEY BEAN CO", 
    nys_product_purchased = NA, 
    sold_to_NYC = 0, 
    beans_processor = 1,
    wholesaler_distributor = 1, 
    nyc_dashboard = NA, 
    lrfs_business = 1, 
    food_product_database = 1, 
    ny = 1, 
    aggregator_cleaner = 1, 
    primary_processor =0, 
    secondary_processor = 0, 
    mwbe = 0, 
    interview_data = 1
  )

#Put in alphabetical order again
df <- df %>% 
  arrange(company_name)

# Add columns we need and put in correct order
df <- df %>% 
  mutate(
    street_address = NA, 
    city = NA, 
    state = NA, 
    zip_code = NA, 
    aggregator_scale = NA, 
    primary_processor_scale = NA,
    secondary_processor_scale = NA,
    max_inv = NA) %>% 
  select(
    company_name, street_address, city, state, zip_code, 
    ny, aggregator_cleaner, primary_processor, 
    secondary_processor,
    aggregator_scale, 
    primary_processor_scale,
    secondary_processor_scale,
    max_inv, 
    lrfs_business,
    mwbe, 
    sold_to_NYC, nys_product_purchased
  )

# Add in zero's for the columns that should be correct
df <- df %>% 
  mutate(across(c(mwbe, sold_to_NYC), 
                ~case_when(
                  is.na(.) ~ 0, 
                  TRUE ~ .))
  )

# Add information on what data means
info <- tibble(
  variable_name = c(
    "company_name", "street_address", "city", "state", 
    "zip_code", "ny", "aggregator_cleaner", "primary_processor" , 
    "secondary_processor", 
    "aggregator_scale", 
    "primary_processor_scale", 
    "secondary_processor_scale",
    "max_inv", 
    "lrfs_business", 
    "mwbe", 
    "sold_to_NYC",
    "nys_product_purchased"), 
  variable_description = c(
    "Company name", 
    "Street address (e.g., 1038 COURT ST)", 
    "City",
    "State abbreviation (e.g., NY)", 
    "Zip code as 5 digits", 
    "0/1 variable indicating if the business is located in NY or not", 
    "0/1 variable indicating if the business is a cleaner/aggregator or not",
    "0/1 variable indicating if the business is a primary processor or not",
    "0/1 variable indicating if the business is a secondary processor or not",
    "Scale of cleaner aggregator business. If not a cleaner/aggregator then leave blank, otherwise fill in with small, medium, large", 
    "Scale of primary processor business. If not a primary processor then leave blank, otherwise fill in with small, medium, large",
    "Scale of secondary processor business. If not a secondary processor then leave blank, otherwise fill in with small, medium, large",
    "Maxiumum inventory capacity for this scale of processor", 
    "0/1 variable indicating if the business is sells NYS source identified product or not",
    "0/1 variable indicating if the business is an MWBE processor or not",
    "0/1 variable indicating if the business is sold to NYC or not",
    "0/1 variable indicating if the business was part of a supply chain that sold NYS source identified product previously to NYC agencies"), 
  notes = c(
    "", 
    "Can be lower case or upper case, fine to just copy and paste from whatever source. Do NOT include state or zip code.", 
    "",
    "Ideally it will be the state abbreviation (NY) but you can also include the full name New York.",
    "",
    "1 indicates yes and 0 indicates no", 
    "1 indicates yes and 0 indicates no", 
    "1 indicates yes and 0 indicates no", 
    "1 indicates yes and 0 indicates no", 
    "Enter the text `small`, `medium`, or `large`. If you have different scale categories then enter these (such as small/large)", 
    "Enter the text `small`, `medium`, or `large`. If you have different scale categories then enter these (such as small/large)", 
    "Enter the text `small`, `medium`, or `large`. If you have different scale categories then enter these (such as small/large)", 
    "Choose on representative business from each category (e.g., small primary processor) and report the max inventory for all busineses in that category",
    "1 indicates yes and 0 indicates no. Please correct if you know the business sells source identified NYS product",
    "1 indicates yes and 0 indicates no. Do not change data in this column. ", 
    "1 indicates yes and 0 indicates no. Do not change data in this column.", 
     "1 indicates yes and 0 indicates no. Do not change data in this column.")
)
    
# Add worksheets
wb <- createWorkbook()
addWorksheet(wb, "data")
writeData(wb, "data", df)

addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save
saveWorkbook(wb, "../data_processed/bean_processor_ADD_DATA.xlsx", 
             overwrite = TRUE)
```

### Processed data
The data that was saved in the step above was sent to the qualitative team and they manually added addresses, indicated those businesses that do not sell beans, added in data on the businesses location along the supply chain where missing (notably if the business was a primary or secondary processor), and estimated values for max inventory capacity. 

We add the wholesale data that are MWBE to the data. 

#### Import and clean

```{r}

# Remove all df's except for the wholesale data of interest
rm(list=setdiff(ls(), c("naics_df", "nyc_wholesale_df", "nyc_bean_max")))

# Import (n= 57)
df <- read_xlsx("../data_processed/bean_processor_PROCESSED.xlsx") %>% 
  clean_names()

# Drop those indicated should be deleted in the notes (n=31)
df <- df %>% 
  mutate(
    delete = case_when(
      str_detect(notes, "Delete") | 
        str_detect(notes, "Too generic of a name.  Not sure what co. it is.") ~ 1, 
      TRUE ~ 0)) %>% 
  select(delete, everything()) %>%
  filter(delete==0)


#Drop if a business is not located in NYS and does not sell NYS identified product (lrfs_business==1) (n=14)
df <- df %>% 
  filter(state=="NY" |
           (state != "NY" & (lrfs_business==1 | mwbe==1)))

# Make unlimited max inventory equal to what NYC purchasaed last year and make numeric
df <- df %>% 
  mutate(
    max_inv_lbs_year = as.numeric(max_inv_lbs_year), 
    max_inv_lbs_year = case_when(
      is.na(max_inv_lbs_year) ~ nyc_bean_max, 
      TRUE ~ max_inv_lbs_year)
    )

# Drop columns we don't need
df <- df %>% 
  select(!c(delete, notes, ends_with("_scale"))) %>% 
  relocate(max_inv_lbs_year, .after = everything())

# Modify wholesale data
naics_df <- naics_df %>% 
  rename(sold_to_nyc = sold_to_NYC)

# Combine aggregator/processor data with final wholesale data
df <- bind_rows(df, naics_df, nyc_wholesale_df)

# Add zeros instead of NA and keep only columns of interest
df <- df %>% 
  relocate(wholesaler_distributor, .after = "ny") %>% 
  mutate(across(wholesaler_distributor:nys_product_purchased, 
                ~case_when(is.na(.) ~ 0, 
                           TRUE ~ .)))

# Make max inventory what NYC purchased for all wholesalers
df <- df %>% 
  mutate(
    max_inv_lbs_year = case_when(
      is.na(max_inv_lbs_year) ~ nyc_bean_max, 
      TRUE ~ max_inv_lbs_year)
    )

# Assume all wholesalers can sell organic and for other supply chain actors only if indicated
df <- df %>% 
  mutate(
    organic = case_when(
      wholesaler_distributor==1 ~ 1, 
      is.na(organic) ~ 0,
      TRUE ~ organic)
  )

```

#### Geocode

```{r}

library(tidygeocoder)

# Add full address
df <- df %>% 
  mutate(
    address = str_c(street_address, ", ", 
                    city, ", ", 
                    state, " ", 
                    zip_code)
  )

# Geocode
df <- df %>%
  geocode(address, 
          method = 'census', 
          lat = latitude, 
          long = longitude,
          full_results = FALSE)

# Manually fix those missing
missing <- df %>% 
  filter(is.na(latitude))

# Fix missing
df <- df %>% 
  mutate(
    latitude = case_when(
      company_name== "COMMUNITY FOOD WORKS (MILESTONE MILL)" ~ 41.937258,
      company_name== "NEW YORK BEAN, LLC" ~ 42.976152,
      company_name== "Unipak" ~ 40.694349,
      company_name== "GO AFRICA GLOBAL LLC" ~ 42.6478356, 
      company_name== "HOME TOWN HAULING & RECYCLING, LLC" ~ 42.6478356, 
      company_name== "MARJORIE MC KEOWN" ~ 42.24757385253906, 
      company_name== "R & N TRANSPORT, INC." ~ 40.692420959472656, 
      company_name== "SOMETHING GOOD TO EAT, INC." ~ 40.79843521118164, 
      company_name == "US FOODS" & 
        street_address=="702 POTENTIAL PKWY SCOTIA" ~ 42.838743, 
      company_name == "US FOODS" & 
        street_address=="95 MAIN ST" ~ 42.889337,
      company_name == "US FOODS" & 
        street_address=="177 NGARA FRNTR FD TERMIN" ~ 42.875127,
      TRUE ~ latitude),
    longitude = case_when(
      company_name== "COMMUNITY FOOD WORKS (MILESTONE MILL)" ~ -74.020285,
      company_name== "NEW YORK BEAN, LLC" ~ -77.870366,
      company_name== "Unipak" ~ -74.297359,
      company_name== "GO AFRICA GLOBAL LLC" ~ -74.7924148, 
      company_name== "HOME TOWN HAULING & RECYCLING, LLC" ~ -74.7924148, 
      company_name== "MARJORIE MC KEOWN" ~ -79.3206558227539, 
      company_name== "R & N TRANSPORT, INC." ~ -73.58958435058594, 
      company_name== "SOMETHING GOOD TO EAT, INC." ~ -73.61048126220703, 
      street_address=="702 POTENTIAL PKWY SCOTIA"  ~ -73.979981,
      company_name == "US FOODS" & 
        street_address=="95 MAIN ST" ~ -73.349132,
      company_name == "US FOODS" & 
        street_address=="177 NGARA FRNTR FD TERMIN" ~ -78.820302,
      TRUE ~ longitude)
    )

# Re-order
df <- df %>% 
  select(latitude, longitude, everything())

rm(missing, naics_df, nyc_wholesale_df)
```

#### Save

Here we save a full version and the version that can be shared publicly with business names and addresses removed. We also drop any columns that are not relevant.

##### Full data

```{r}

# Save full file - not to be shared publicly
write_csv(df, "../data_final/beans_processor_distributor_full.csv")

```

##### Final data

Replace business names with an unique id. We also drop beans_processor as it was replaced with primary and secondary and is confusing to include both. Lastly, we drop variables related to scale and local ownership as most of the data is missing and we are relying on other sources.

```{r}

# Drop columns not needed
df <- df %>% 
  select(
    address, company_name,
    latitude, longitude,
    ny, wholesaler_distributor, aggregator_cleaner, 
    primary_processor, secondary_processor,
    lrfs_business, mwbe, sold_to_nyc, 
    organic, max_inv_lbs_year)

# Add a unique identifier for business name
df <- df %>%
  mutate(id = row_number() + 9999) %>% 
  select(id, everything())

# create a file with business name/address and identifier and save
beans_business_id <- df %>% 
  select(id, address, company_name)

write_csv(df, "../data_processed/beans_business_id.csv")

# Drop business name/address from file
df <- df %>% 
  select(-address, -company_name)

# Save to the folder with all data
write_csv(df, "../data_final/beans_processor_distributor.csv")

# Save to the public data folder - no max inventory
df_nomaxinv <- df %>% 
  select(-max_inv_lbs_year)

write_csv(df_nomaxinv, "/Users/alg24/Library/CloudStorage/Box-Box/Research/CityFoodPolicy/2023.NYC_FFARRockefeller/research/data_master_record/CityFoodPolicy/drybeans/data_final/beans_processor_distributor.csv")

```
