---
format: html
editor: visual
---

```{r}
#| include: false
# When included in the book, working dir is drybeans/ not data_inputs_code/
if (!dir.exists("../data_processed") && dir.exists("data_inputs_code")) {
  knitr::opts_knit$set(root.dir = normalizePath("data_inputs_code"))
}
```

```{r}
library(tidyverse)
library(readxl)
```

We estimate cost of production for dry bean and soybean producers using a combination of data sources including the restricted access Census of Agriculture, crop budgets from select states, and interview data. We get the number of farms and harvested acres by county for dry bean and soybean producers in New York using the public Census of Agriculture data.

### Cost of production {#sec-drybeans-costofproduction}

[Final data: Cost of production for dry beans and soybeans.](https://github.com/FoodSystemsModeling/CityFoodPolicy/tree/main/drybeans/data_final/production_cost.xlsx)

Using the restricted access data from the Census of Agriculture, we use the NASS defined variable, total expense, which includes all production expenses listed in the production expense section of the census and includes both fixed and variable expenses. Data are pulled for producers with dry bean sales. However, these data are at the whole farm level. We do not know the cost of production of beans specifically.

#### Dry beans

To understand cost of production for New York bean producers, we use data from interviews conducted of New York bean producers, which estimate a cost of production of \$525 per acre.

We also need a cost of production for bean producers outside of New York, as those are the producers currently supplying the New York City beans. Based on discussions with Michael, most of the beans come from North Dakota, so we base our cost of production on an enterprise budget from North Dakota. We also include the other crop budgets we could find for comparison/to provide range:

- [North Dakota, 2024](https://www.ndsu.edu/agriculture/sites/default/files/2024-02/2024%20North%20East%20ND%20Projected%20Crop%20Budget%20-%20PDF.pdf)
  - Cost per acre: \$434.05
- [Wyoming](https://wyoextension.org/publications/html/B1315-2/)
  - Cost per acre: \$598.70, cost per CWT \$23.95
- [Colorado](https://abm.extension.colostate.edu/enterprise-budgets-crop/)
  - Cost per acre: \$454.57, cost per CWT \$113.64
- [Texas](https://aggie-horticulture.tamu.edu/smallacreage/crops-guides/vegtables/dry-beans/)
  - Cost per acre: \$400

**Conclusions** We will use the estimated cost of production data for New York dry bean producers combined with Census of Agriculture data to estimate cost of production for different types of operations. For those outside of NY, we will provide a range based on the enterprise budgets we were able to find.

From the Census of Agriculture, we have an average dry bean yield in NY of 20.5 CWT per acre to convert to production cost per lb.

The New York cost of production per acre is based on June's assumption.

We increase/decrease the average cost of production by scale and MWBE status based on U.S. data from the Census of Agriculture. We calculate the percent difference of each category from all producers to scale data. Data is not available for all categories in New York due to small sample sizes.

```{r}
library(tidyverse)
library(readxl)
library(openxlsx)
library(janitor)

# Average cost of production based on interviews
ny_production_cost_acre <- 525

# Standard deviation based on the ratios from the highest and lowest in crop budgets
ny_production_cost_acre_sd <- (598.70-400)/4

# Import US data
drybeans <- read_xlsx("../data_processed/US data/03.beans_final_results_suppress.xlsx", 
                      sheet = "summary_organic")

# Keep cost of production per acre only and delete unneeded data
drybeans <- drybeans %>% 
  filter(variable_name=="exp_acre" & 
           crop == "dry_beans") %>% 
  select(weighted_n, sales_class_agg, mwbe_fct, organic_fct, statistic, value)

# Pivot wider so mean/se in separate columns
drybeans <- drybeans %>% 
  pivot_wider(
    names_from = statistic, 
    values_from = value
  )

# Add standard deviation
drybeans <- drybeans %>% 
  mutate(
    sd = se * sqrt(weighted_n)
  ) %>% 
  select(-se, -weighted_n)

# Add in totals for the U.S. aggregated 
drybeans <- drybeans %>% 
  add_row(
    sales_class_agg="All", 
    mwbe_fct = "All", 
    organic_fct = "Not organic", 
    mean = 950) %>%
  add_row(
    sales_class_agg="All", 
    mwbe_fct = "All", 
    organic_fct = "Organic", 
    mean = 1250) 

# Compute scalers for organic and not 
organic_scaler <- drybeans %>% 
  filter(sales_class_agg=="All" & 
           organic_fct=="Organic") %>% 
  pull(mean)

conventional_scaler <- drybeans %>% 
  filter(sales_class_agg=="All" & 
           organic_fct=="Not organic") %>% 
  pull(mean)

# Drop all data
drybeans <- drybeans %>% 
  filter(sales_class_agg !="All")

# Scale data
drybeans <- drybeans %>% 
  mutate(
    percent_diff_mean = case_when(
      organic_fct=="Organic sales" ~ (mean - organic_scaler)/organic_scaler, 
      organic_fct=="No organic sales" ~ (mean - conventional_scaler)/conventional_scaler)
  )


# Add production cost per acre for beans from crop budget
drybeans <- drybeans %>% 
  mutate(
    year = 2024,
    baseline = ny_production_cost_acre, 
    baseline_sd = ny_production_cost_acre_sd,
    production_cost_acre_mean = baseline * (1 + percent_diff_mean),
    production_cost_acre_sd   = baseline_sd * (1 + percent_diff_mean)
    )

# Keep only data we need
drybeans <- drybeans %>% 
  select(sales_class_agg, mwbe_fct, 
         organic_fct, 
         production_cost_acre_mean, production_cost_acre_sd)


# Add in $ per hectare (1 hectare) = 2.471 acres)
drybeans <- drybeans %>% 
  mutate(
    production_cost_hectare_mean = production_cost_acre_mean*2.471, 
    production_cost_hectare_sd = production_cost_acre_sd*2.471
  )

# Add crop
drybeans <- drybeans %>% 
  mutate(
    crop = "drybeans"
  ) %>% 
  select(crop, everything())

```

#### Soybeans

We use Census of Agriculture data and keep those producers where total sales is equivalent to soybean sales. Because we have soybean producers only, then the total cost per acre is for soybeans only and we don't need to use a crop budget. We do pull crop budgets for reference.

In NY there are not enough farms in many categories. So we modify that NY data that we have up or down based on the national numbers. There are not enough organic soy producers in New York, so we assume the same difference between conventional and organic in New York as we see nationally.

If data are missing, that means there are no farms nationally in that category (i.e., only soybeans, organic, over \$1M). We assume all organic production expenses are increased by the amount that organic is higher at the national level for all soybean only producers.

Crop budgets for comparison:

- [Ohio in 2022](https://farmoffice.osu.edu/farm-management/enterprise-budgets#2024)
  - Conventional: 761/acre.
  - There is no data for organic until 2024, which is 946/acre. For comparison, conventional soy in 2024 was 754/acre.
- [Louisiana in 2022](https://www.lsuagcenter.com/articles/page1641147769855)
  - 454/acre
- [Alabama in 2025](https://www.aces.edu/blog/topics/farm-management/soybean-enterprise-budget/)
  - 471/acre

```{r}

# Import data
soybeans <- read_xlsx(
  "../data_processed/US data/03.beans_final_results_suppress.xlsx", 
  sheet = "soybeans_production_cost")

# Keep relevant columns
soybeans <- soybeans %>% 
  select(sample, weighted_n, sales_class_agg, mwbe_fct, 
         organic_fct, statistic, value)

# Drop organic data
soybeans <- soybeans %>% 
  filter(
    (!is.na(sales_class_agg) & 
           organic_fct == "No organic sales") | 
  is.na(sales_class_agg)
  )
  
# Pivot wider so mean/sd in separate columns
soybeans <- soybeans %>% 
  pivot_wider(
    names_from = statistic, 
    values_from = value
  )

# Convert se to sd
soybeans <- soybeans %>% 
  mutate(sd = se * sqrt(weighted_n))

# Compute percent difference in mean and sd between NY and US (conventional)
pct_diff_conventional <- soybeans %>% 
  filter(organic_fct == "No organic sales" & is.na(sales_class_agg)) %>% 
  summarise(
    pct_diff_mean = (mean[sample == "New York"] - mean[sample == "US"]) / mean[sample == "US"]
    ) %>% 
  pull(pct_diff_mean)

# Compute percent diff for US between organic and conventional
pct_diff_organic <- soybeans %>% 
  filter(sample == "US" & is.na(sales_class_agg)) %>% 
  summarise(
    pct_diff_mean = (mean[organic_fct == "Organic sales"] - mean[organic_fct == "No organic sales"]) / mean[organic_fct == "No organic sales"]
  ) %>% 
  pull(pct_diff_mean)

# Drop rows with aggregate data
soybeans <- soybeans %>% 
  filter(!is.na(sales_class_agg)) %>% 
  select(-sample, -weighted_n)

# Compute the estimated mean for conventional NY producers by scaling US subgroup means
soybeans <- soybeans %>%
  mutate(
    production_cost_acre_mean  = mean * (1 + pct_diff_conventional),
    production_cost_acre_sd  = sd * (1 + pct_diff_conventional)
)
    
# Add rows for organic, scaled based on national numbers
organic <- soybeans %>% 
  mutate(
    production_cost_acre_mean  = production_cost_acre_mean *(1 + pct_diff_organic), 
    production_cost_acre_sd  = production_cost_acre_sd *(1 + pct_diff_organic), 
    organic_fct = "Organic sales"
  )

# Combine
soybeans <- bind_rows(soybeans, organic)
rm(organic)

# Keep columns of interest
soybeans <- soybeans %>% 
  select(-mean, -se, -sd) 

# Rename and add in $ per hectare (1 hectare) = 2.471 acres)
soybeans <- soybeans %>% 
  mutate(
    production_cost_hectare_mean = production_cost_acre_mean*2.471,
    production_cost_hectare_sd = production_cost_acre_sd*2.471)

# Add crop
soybeans <- soybeans %>% 
  mutate(
    crop = "soybeans"
  ) %>% 
  select(crop, everything())
```

#### Save

Combine all data into one file and save. [Final cost of production data](https://github.com/FoodSystemsModeling/CityFoodPolicy/tree/main/drybeans/data_final/production_cost.xlsx)

```{r}

# Combine dry bean and soybean data
production_cost <- bind_rows(drybeans, soybeans) 

# Write to file
wb <- createWorkbook()

# Add info
info <- tibble(
  sheet_name = "production_cost", 
  description = "Mean and standard deviation for cost of production per acre and hectare for dry bean and soybean growers. Costs are by crop, scale (under/over $M), organic, and MWBE).",
  source = "Restricted access Census of Agriculture and crop budgets. See Quarto file for more information."
)

# Add worksheets and save
addWorksheet(wb, "info")
writeData(wb, "info", info)

addWorksheet(wb, "production_cost")
writeData(wb, "production_cost", production_cost)

saveWorkbook(wb, "../data_final/production_cost.xlsx", 
             overwrite = TRUE)

rm(list=ls())
```

### County production

[Final data: Number of farms and acres of dry beans and soybeans in New York.](https://github.com/FoodSystemsModeling/CityFoodPolicy/blob/main/drybeans/data_final/acres_farms_county.xlsx)

We use the number of farms and harvested acres by county for dry bean and soybean producers in New York to ensure our model has the correct number of farms and acres.

We download 2022 data from [U.S. Department of Agriculture, National Agricultural Statistics Service (USDA NASS), Data, Large Datasets](https://www.nass.usda.gov/datasets/).

We keep state totals and county-level data for number of farms, acres harvested, and quantity harvested for drybeans (DRY EDIBLE BEANS, EXCLUDING CHICKPEAS AND LIMAS (CWT)) and soybeans (SOYBEANS FOR BEANS (BUSHELS)).

In counties where data is missing due to disclosure rules, we allocate remaining acres across those counties. After cleaning, if all rows are NA, then there are no farms in the county, if the number of farms is positive but NAs for the other columns then the data was not disclosed.

#### Import and clean

```{r}
wb <- createWorkbook()

# Import county-level data for NY
coa <- read_tsv("../data_raw/qs.census2022.txt") %>% 
  clean_names()

# Keep NY only, state and county data 
coa <- coa %>% 
  filter(state_alpha=="NY")

# Keep data of interest
varlist <- c(
  "BEANS, DRY EDIBLE, (EXCL CHICKPEAS & LIMA) - OPERATIONS WITH AREA HARVESTED",
  "BEANS, DRY EDIBLE, (EXCL CHICKPEAS & LIMA) - ACRES HARVESTED",
  "BEANS, DRY EDIBLE, (EXCL CHICKPEAS & LIMA) - PRODUCTION, MEASURED IN CWT",
  "SOYBEANS - OPERATIONS WITH AREA HARVESTED",
  "SOYBEANS - ACRES HARVESTED", 
  "SOYBEANS - PRODUCTION, MEASURED IN BU"
)

coa <- coa %>% 
  filter(short_desc %in% varlist &
           domain_desc == "TOTAL")

# Add fips
coa <- coa %>% 
  mutate(fips = case_when(
    agg_level_desc=="STATE" ~ state_fips_code, 
    agg_level_desc=="COUNTY" ~ str_c(state_fips_code, 
                                     county_code, sep = ""))
  )

# Make county name lowercase
coa <- coa %>% 
  mutate(
    county_name = str_to_title(county_name)
  )

# Convert value to numeric
coa <- coa %>% 
  mutate(value = as.numeric(str_remove_all(value, ",")))

# Define crop (soy/dry)
coa <- coa %>% 
  mutate(
    crop = case_when(
      str_detect(short_desc, "SOYBEANS") ~ "soybeans", 
      TRUE ~ "drybeans"
    )
  )

# Select columns of interest
coa <- coa %>% 
  select(fips, county_name, crop, short_desc, unit_desc, value) 

# Pivot wider
coa <- coa %>% 
  select(-short_desc) %>%
  pivot_wider(
    names_from = c(crop, unit_desc), 
    values_from = value
  ) %>% 
  select(fips, county_name, starts_with("drybeans"), starts_with("soybeans"))

# Get state totals
coa_state <- coa %>% 
  filter(is.na(county_name)) %>% 
  select(-fips, -county_name) %>% 
  rename_with(~paste0(., "_NYtotal"))

# Drop state data
coa <- coa %>% 
  filter(fips != "36")

# Add a dummy for western counties
western_counties <- c("Allegany", "Cattaraugus",
                         "Chautauqua", "Erie",
                         "Genesee", "Livingston",
                         "Monroe", "Niagara",
                         "Ontario", "Orleans",
                         "Steuben", "Wyoming", "Yates") 
coa <- coa %>% 
  mutate(
    west = case_when(
      county_name %in% western_counties ~ 1, 
      TRUE ~ 0)
  )

```

#### Allocate missing acres

Western counties have larger farms than other counties in NY. We take the average farm size in Western counties divided by the average size in other counties for each commodity and use the number of farms plus if located in a western county to allocate non-disclosed acres.

##### Drybeans

Western counties have larger farms than other counties in NY. We take the average farm size in Western counties divided by the average size in other counties. Dry bean farms are about 3 time larger in Western counties than in the rest of New York.

```{r}

# Make separate dfs for dry and soy
dry_coa <- coa %>% 
  select(fips, county_name, west, drybeans_OPERATIONS, 
         drybeans_ACRES
         )

# Scaling - average size in West 3x larger than rest of NY
scaling <- dry_coa %>% 
  mutate(
    ave_acres = drybeans_ACRES/drybeans_OPERATIONS,
    ) %>% 
  group_by(west) %>% 
  summarise(ave_acres = mean(ave_acres, na.rm = TRUE)) %>% 
  mutate(
    scaling_ratio = ave_acres / ave_acres[west == 0]
  ) %>% 
  filter(west==1) %>%
  pull(scaling_ratio)

# Drop counties with no farms
dry_coa <- dry_coa %>% 
  filter(!is.na(drybeans_OPERATIONS))

# Define total acres in NY
total_acres <- coa_state %>%
  select(drybeans_ACRES_NYtotal) %>% 
  pull()

# Number of acres not disclosed
undisclosed_acres <- total_acres - dry_coa %>% 
  summarise(sum(drybeans_ACRES, na.rm = TRUE)) %>%
  pull()

# Allocate acres to West and rest of NY using scaling ratio
west_acres <- undisclosed_acres * (scaling / (scaling + 1))
nonwest_acres <- undisclosed_acres * (1 / (scaling + 1))

# Number of farms with acres not disclosed
undisclosed_farms <- dry_coa %>% 
  group_by(west) %>%
  filter(is.na(drybeans_ACRES)) %>% 
  summarise(undisclosed_farms = sum(drybeans_OPERATIONS, na.rm = TRUE)) %>% 
  ungroup()

west_farms <- undisclosed_farms %>% 
  filter(west==1) %>% 
  select(undisclosed_farms) %>%
  pull()

nonwest_farms <- undisclosed_farms %>% 
  filter(west==0) %>% 
  select(undisclosed_farms) %>%
  pull()

# Allocate remaining acres, with ~3x more acres in Western states with missing values
dry_coa <- dry_coa %>% 
  mutate(
    drybeans_ACRES = case_when(
      !is.na(drybeans_ACRES) ~ drybeans_ACRES, 
      is.na(drybeans_ACRES) & west==1 ~ west_acres*(drybeans_OPERATIONS/west_farms), 
      is.na(drybeans_ACRES) & west==0 ~ nonwest_acres*(drybeans_OPERATIONS/nonwest_farms)
    )
  )
  
```

##### Soybeans

For soybeans, all non-disclosed farms are not located in the west. We allocate acres evenly assuming an average sized farm.

```{r}

# Make separate dfs for dry and soy
soy_coa <- coa %>% 
  select(fips, county_name, west, soybeans_OPERATIONS, 
         soybeans_ACRES)

# Drop counties with no farms
soy_coa <- soy_coa %>% 
  filter(!is.na(soybeans_OPERATIONS))

# Define total acres in NY
total_acres <- coa_state %>%
  select(soybeans_ACRES_NYtotal) %>% 
  pull()

# Number of acres not disclosed
undisclosed_acres <- total_acres - soy_coa %>% 
  summarise(sum(soybeans_ACRES, na.rm = TRUE)) %>% 
  pull()

# Number of farms with acres not disclosed
undisclosed_farms <- soy_coa %>% 
  filter(is.na(soybeans_ACRES)) %>% 
  summarise(sum(soybeans_OPERATIONS)) %>% 
  pull()

# Number of farms with acres not disclosed
undisclosed_farms <- soy_coa %>% 
  filter(is.na(soybeans_ACRES)) %>% 
  summarise(undisclosed_farms = sum(soybeans_OPERATIONS, na.rm = TRUE)) %>%
  pull()

# Allocate remaining acres, assuming average size (none are western)
soy_coa <- soy_coa %>% 
  mutate(
    soybeans_ACRES = case_when(
      !is.na(soybeans_ACRES) ~ soybeans_ACRES, 
      is.na(soybeans_ACRES) ~ undisclosed_acres*(soybeans_OPERATIONS/undisclosed_farms)
    )
  )
  
```

##### Proportion by group

Get the proportion of farms in each group so they can be allocated across counties, using data from the restricted access census. 

```{r}

n <- read_xlsx("../data_processed/2026.07.15/CFP.04.beans_final_results_suppress.xlsx", 
               sheet = "summary")


# get sample sizes by group and crop
n <- n %>%
  group_by(crop, sales_class_agg, mwbe_fct) %>% 
  distinct(weighted_n)

# Get total sample size by crop
n <- n %>% 
  group_by(crop) %>%
  mutate(
    total = sum(weighted_n)
    ) %>% 
  ungroup()

# Caluculate ratios
n <- n %>% 
  mutate(
    ratio = weighted_n/total
  )

# Add worksheet 
addWorksheet(wb, "ratios_by_group")
writeData(wb, "ratios_by_group", n)

```

#### Join and save

```{r}

# Join
coa <- dry_coa %>% full_join(soy_coa)

# Drop west
coa <- coa %>% 
  select(-west)

# Fill NAs with 0
coa <- coa %>% 
  mutate(across(c(drybeans_OPERATIONS, drybeans_ACRES), 
         ~case_when(is.na(.) ~ 0, 
                    TRUE ~ .))
  )

# Add worksheet
addWorksheet(wb, "acres_farms")
writeData(wb, "acres_farms", coa)

# Add info
info <- tibble(
    sheet_name = c(
      "acres_farms", "ratio_by_group"),
    sheet_description = c(
      "Number of acres and farms by county for soybean and drybean operations in New York", 
      "Ratio of farms by group, used to allocate total number of farms into groups based on MWBE status and scale"),
    source = c(
      "2022 Census of Agriculture, public data, Table 25. Field Crops: 2022 and 2017", 
      "Restricted access 2022 Census of Agriculture"),
    notes = c(
      "We allocate all missing acres acros counties with acres not disclosed based on the number of farms in the county. We allocate more acres to farms in the western part of the state based on the average size of farms in the west vs. rest of New York. Dry bean farms are around 3 times bigger in the west and soybean around 1.2 times larger.", 
      NA)
  )

# Add worksheet
addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save as a separate finfo# Save as a separate file 
saveWorkbook(wb, "../data_final/acres_farms_county.xlsx", 
             overwrite = TRUE)

```

### Differentiated market

[Final data: Total demand for New York State differentiated dry beans.](https://github.com/FoodSystemsModeling/CityFoodPolicy/blob/main/drybeans/data_final/differentiated_total_demand.xls)

To determine to total market for NYS differentiated beans, four dry bean growers with food grade cleaning capacity who sell directly to buyers were contacted (interviews, emails, phone calls). Sales data was provided by the producer. (Numbers were confirmed to make sure they are not double counted in the totals).

```{r}
wb <- createWorkbook()

# Create data frame with information
diff <- tibble(
  type = "differentiated", 
  commodity = "drybeans", 
  variable = "total_demand_kg", 
  value = 1248762/2.205
)

## Add worksheet
addWorksheet(wb, "differentiated_market_size")
writeData(wb, "differentiated_market_size", diff)

## Info
info <- tibble(
  sheet_name = "differentiated_market_size", 
  description = "Total size of the differentiated market (i.e., NYS source identified) dry beans", 
  source = "interviews", 
  notes = "Data are reported in total killograms of dry beans"
)

addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save
saveWorkbook(wb, "../data_final/differentiated_total_demand.xlsx", 
             overwrite = TRUE)

```

### Farmer characteristics

We use data from the restricted access 2022 Census of Agriculture to paramertize agents in the model. [Exported data](https://github.com/FoodSystemsModeling/CityFoodPolicy/blob/main/drybeans/data_processed/2026.07.15/CFP.04.beans_final_results_suppress.xlsx) and [code](https://github.com/FoodSystemsModeling/CityFoodPolicy/blob/main/drybeans/data_processed/2026.07.15/CFP.03.beans_final_results_approved.qmd) used to generate results.

Sample includes farms in the Northeast that are (1) dry bean producers and (2) soybean producers that don't produce dry beans and are not primary commodity livestock/dairy.

The sheet titled "summary" includes summary statistics (mean/standard error) for a variety of different farm-level characteristics separately for dry bean and soybean producers by scale, MWBE, and organic.

The sheet titled "summary_landuse_organic" includes summary statistics specifically related to land use for dry bean and soybean producers for those with organic sales and those without.

The sheet titled "metdata" contains definitions of all variables and the sheet titled "info" describes the data at a high level along with the sample.

Notes: Operations with less than 1,000 in sales are dropped, if N\<5 it is not included, (D) indicated not included due to disclosure risk.
