---
title: "Lettuce"
format: html
editor: visual
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE, 
                      warning = FALSE)
```

In this file we use 2022 Census of Agriculture data to understand the expenditure patterns and land use of cabbage and leafy green farms in New York.

## Import data

```{r}
library(tidyverse)
library(srvyr)
library(openxlsx)
library(broom)

# Load data
load("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/data_processed/coa_22.RData")

# Import my functions 
source("P:/pr-usda-nass-jablonski-csu23-coaarmslf/general/myfunctions.R")

# Begin excel output file 
wb <- createWorkbook()

# define data frame for the northeast
df <- coa_22 %>% 
  filter(state_name == "New York")

rm(coa_22)

```

## Define variables

In the project, leafy greens are defined as the leaves of green vegetables such as romaine lettuce, spinach, kale, collard greens, cabbage, watercress, arugula, bok choy, beet greens, and Swiss chard, when eaten as food. Separately, we are interested in cabbage.

We define leafy green operations as those with positive harvested acres of:

-   K389 Cabbage, Chinese, Harvested, Tenth-Acres
-   K407 Collards Harvested, Tenth-Acres
-   K425 Kale Harvested, Tenth-Acres
-   K428 Lettuce, Head, Harvested, Tenth-Acres
-   K430 Lettuce, Leaf, Harvested, Tenth-Acres
-   K432 Lettuce, Romaine, Harvested, Tenth-Acres
-   K431 Mustard Greens Harvested, Tenth-Acres
-   K457 Spinach Harvested, Tenth-Acres


We define cabbage operations as those with positive harvested acres for cabbage head (k391). We also know operations with cabbage head harvested for fresh market (k1991) and for processing (k2391)

We do not want to include operations that are greenhouse only. Operations with more than 1/2 acre production are very unlikely to have high tunnels and we assume they are a greenhouse operation. There are 21780 square feet in an 1/2 acre.

We separate operations by organic and not organic.

Notes:

-   k2503 - Vegetables and fresh cut herbs, Sales is from section 11 and only includes crops grown under protection.
-   K1374 Total Vegetable, Melons, and Potatoes Sales

```{r}

# Define operations with vegetable sales 
df <- df %>% 
  mutate(
    veg_glass_sqft = k503,
    veg_glass_sales = k2503,
    greenhouse_cat = case_when(
      veg_glass_sqft> 21780 ~ 1, 
      TRUE ~ 0),
    leafy_green_10thacres = k389 + k407 + k425 + k428 + k430 + k432 + k431 + k457,
    leafy_green_farm = case_when(
      leafy_green_10thacres>0 ~ 1, 
    TRUE ~0),
    cabbage_10thacres = k391, 
    cabbage_farm = case_when(
      cabbage_10thacres>0 ~ 1, 
      TRUE ~ 0), 
    veg_sales = k1347)


# Define production expenses grouped and ungrouped
df <- df %>% 
  mutate(
    labor_cwork = labor + cwork, 
    labor_cwork_var_pct = labor_cwork/totvarexp, 
    rent_land_buildings = k1513, 
    rent_lease_machinery = k1514, 
    rent_land_buildings_fixed_pct = k1513/totfixedexp,
    rent_lease_machinery_fixed_pct = k1514/totfixedexp)

# Define depreciation as a percent of total machinery and equipment 
df <- df %>% 
  mutate(
    depreciation = k1520, 
    depreciation_machinery_pct = depreciation/k943)

```

## Final data frame

```{r}

# Keep only operations with sales in these categories 
df <- df %>% 
  filter(leafy_green_farm==1 |
           cabbage_farm==1)

# Drop greenhouse only operations
df <- df %>% 
  filter(greenhouse_cat==0)

```

## Summary stats - total

```{r}

varlist <- "tvp"

sum1 <- df %>% 
  filter(leafy_green_farm==1) %>%
  mutate(
    groupbyvar = organic_farm) %>% 
  as_survey(weight= weight) %>% 
  summary_total_continous() %>% 
  mutate(farm_type = "leafy_green_farm")

sum2 <- df %>% 
  filter(cabbage_farm==1) %>%
  mutate(
    groupbyvar = organic_farm) %>% 
  as_survey(weight= weight) %>% 
  summary_total_continous() %>% 
  mutate(farm_type = "cabbage_farm")

sum <- bind_rows(sum1, sum2)

rm(sum1, sum2)

# Make 4 significant digits
sum <- sum %>% 
  mutate(value = signif(value, 4))

# Add Worksheet
addWorksheet(wb, "sum")
writeData(wb, "sum", sum)
```

## OLS

```{r}

## Define varlist
yvar <- df$variables %>% 
  select(
    gcfi_0_75, gcfi_75_350, gcfi_350_1M, gcfi_over_1M,
    tvp, gcfi, veg_sales, roa, opm, infi, operated_acres,
    fert_chemical, seedp, animalpurchase, feed, 
    fuelo, util, mainr, hired_labor, contract_labor, cwork,
    labor, totvarexp,
    rent_lease, 
    rent_land_buildings, rent_lease_machinery, 
    property_tax, interest, veterinary, otherexp, 
    totfixedexp,
    ends_with("_var_pct"), ends_with("_fixed_pct"), 
    depreciation, depreciation_machinery_pct,
    fullowner, partowner, fulltenant, 
    crop_insurance, gov_payments_cat, 
    leafy_green_10thacres, cabbage_10thacres, 
    cover_crop, reduced_till, no_till,
    beginning, fieldcrop, fruitveg, othercrops, 
    livestock, dairy, poultry, otheranimal) %>% 
  names()

yvar <- yvar[!grepl("alt_var_pct$", yvar)]
yvar <- yvar[!grepl("alt_fixed_pct$", yvar)]

xvar <- "organic_farm"

# OLS to loop through yvars
m <- map(yvar, function(yvar){
  model <- df$variables %>% 
    filter(leafy_green_farm==1) %>%
    select(weight, all_of(xvar), all_of(yvar)) %>% 
    lm(paste(yvar, " ~ ", xvar),
       weight = weight, 
      data = .)
  list(
    tidied = tidy(model) %>% 
      mutate(yvar = yvar), 
    glanced = glance(model) %>% 
      mutate(yvar = yvar)
  )
})

tidy <- bind_rows(m) %>%
  select(starts_with("tidied")) %>%
  unnest(cols = c(tidied)) %>% 
  mutate(farm_type = "leafy_green_farm")

glance <- bind_rows(m) %>%
  select(starts_with("glanced")) %>%
  unnest(cols = c(glanced)) %>% 
  mutate(farm_type = "leafy_green_farm")
rm(m)

m <- map(yvar, function(yvar){
  model <- df$variables %>% 
    filter(cabbage_farm==1) %>%
    select(weight, all_of(xvar), all_of(yvar)) %>% 
    lm(paste(yvar, " ~ ", xvar),
       weight = weight, 
      data = .)
  list(
    tidied = tidy(model) %>% 
      mutate(yvar = yvar), 
    glanced = glance(model) %>% 
      mutate(yvar = yvar)
  )
})

tidy_2 <- bind_rows(m) %>%
  select(starts_with("tidied")) %>%
  unnest(cols = c(tidied)) %>% 
  mutate(farm_type = "cabbage_farm")

glance_2 <- bind_rows(m) %>%
  select(starts_with("glanced")) %>%
  unnest(cols = c(glanced)) %>% 
  mutate(farm_type = "cabbage_farm")

# bind 
tidy <- bind_rows(tidy, tidy_2)
glance <- bind_rows(glance, glance_2)

# Write to file 
addWorksheet(wb, "ols_tidy")
writeData(wb, sheet = "ols_tidy", tidy)

addWorksheet(wb, "ols_glance")
writeData(wb, sheet = "ols_glance", glance)

rm(m, tidy, glance, tidy_2, glance_2)

```

## Save data

```{r}

# Add info
info <- tibble(
  sheet_name = c("ols_tidy", "ols_glance"), 
  description = c("Model results from a linear regression",
                  "Model statistics from a linear regression (including number of observations, nobs"),
  sample = "Leafy green operations in the New York")


# Add worksheet
addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save
saveWorkbook(wb, "data_final/lettuce.xlsx", overwrite = TRUE)
save(wb, file = "data_final/lettuce.RData")

```
