---
title: "Leafy Greens Focus Group, Tables and Figures"
format: 
  html:
    toc: true
    number-sections: true
  docx:
    toc: true
    number-sections: true
editor: visual
prefer-html: false
execute: 
  warning: false
  message: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

# Data Description

We use data compiled from the 2022 restricted access farm-level data from the Census of Agriculture. All dollars are reported in 2022 dollars. 

We include two operations with leafy green or cabbage production in New York. Where cabbage operations are defined as those with positive harvested acres for cabbage head and leafy green operations are defined as those operations with positive harvested acres of:

-   Cabbage, Chinese, Harvested, Tenth-Acres
-   Collards Harvested, Tenth-Acres
-   Kale Harvested, Tenth-Acres
-   Lettuce, Head, Harvested, Tenth-Acres
-   Lettuce, Leaf, Harvested, Tenth-Acres
-   Lettuce, Romaine, Harvested, Tenth-Acres
-   Mustard Greens Harvested, Tenth-Acres
-   Spinach Harvested, Tenth-Acres

Data are grouped by operations with certified organic production and those with conventional production.

Tables:

-   Average producer summary statistics
-   Variable expenses
-   Production management practices

Notes: We do not want to include operations that are greenhouse only. Operations with more than 1/2 acre production are very unlikely to have high tunnels and we assume they are a greenhouse operation, these operations are not included in the analysis.

```{r}

library(tidyverse)
library(readxl)
library(knitr)
library(kableExtra)
library(scales)

# Turn off scientific notation 
options(scipen = 999)

# Pull data defined in data cleaning code
invisible(knitr::purl("01.greens_COA_datacleaning.qmd", 
            output = "01.greens_COA_datacleaning.R"))
source("01.greens_COA_datacleaning.R")

```

\newpage

# Average producer summary statistics

-   Gross cash farm income (GCFI) (includes total sales and government payments)
-   Total sales
-   Total vegetable sales
-   Total acres operated
-   10th acre harvested for leafy greens and cabbage
-   Primary commodity
-   Land tenure (full owner, part owner, full tenant)
-   Beginning (any operator has 10 years or less experience)

```{r}

# Select variables of interest 
varlist <- c("gcfi", "tvp", "veg_sales", 
              "operated_acres", 
              "leafy_green_10thacres", 
              "cabbage_10thacres", 
             "livestock", "fieldcrop", "dairy", 
             "poultry", "othercrops", "fruitveg", "otheranimal", 
             "crop_insurance", "gov_payments_cat", 
             "fullowner", "fulltenant", "partowner", 
             "beginning")

# Create data frame with variables of interest keeping mean only
sum <- df %>% 
  filter(yvar %in% varlist) %>% 
  select(farm_type, short_variable_description, 
         yvar, mean_conventional, mean_organic) %>% 
  rename(conventional = mean_conventional, 
         organic = mean_organic)

# Format table
sum <- sum %>% 
  mutate(across(c(conventional, organic), 
                ~case_when(
                  . > 1 | yvar=="cabbage_10thacres" ~ 
                    prettyNum(., big.mark = ",", digits = 2),
                  TRUE ~ percent(., accuracy = 1.0))))

sum_greens <- sum %>%
  filter(farm_type=="leafy_green_farm") %>% 
  select(!farm_type) %>%
  rename(`Conventional (n=609)` = conventional, 
         `Organic (n=172)` = organic, 
         Variable = short_variable_description) %>% 
  select(!yvar)

sum_cabbage <- sum %>%
  filter(farm_type=="cabbage_farm") %>% 
  select(!farm_type) %>%
  rename(`Conventional (n=386)` = conventional, 
         `Organic (n=111)` = organic, 
         Variable = short_variable_description) %>% 
  select(!yvar)
```

## Average producer summary statistics- Leafy green farmers

```{r sum_greens}

kable(sum_greens) %>% 
  kable_styling() %>% 
  footnote(general = "All dollar amounts are in 2022 dollars")
```

## Average producer summary statistics- Cabbage farmers

```{r sum_cabbage}

kable(sum_cabbage) %>% 
  kable_styling() %>% 
  footnote(general = "All dollar amounts are in 2022 dollars")
```

\newpage

## Percent of New York leafy green operations by scale (gross cash farm income)

```{r gcfi_bar_greens, fig.align="center", fig.width=7, fig.height=6, out.width="100%"}

rm(sum, sum_cabbage, sum_greens)

## GCFI stacked bar chart 
# Format data 
gcfi_cat <- df %>% 
  select(!c(starts_with("se"), alt_short_variable_description)) %>%
  filter(str_detect(yvar, "gcfi_") & 
           farm_type=="leafy_green_farm") %>%
  pivot_longer(
    cols = !c(farm_type, short_variable_description, yvar),
    names_to = "variable", 
    values_to = "value") %>% 
  mutate(
    groupbyvar = case_when(
      variable == "mean_organic" & farm_type == "leafy_green_farm" ~ 
        "Organic\n(n=172)",
       variable == "mean_conventional" & farm_type == "leafy_green_farm" ~ 
        "Conventional\n(n=609)", 
      variable == "mean_organic" & farm_type == "cabbage_farm" ~ 
        "Organic\n(n=111)",
       variable == "mean_conventional" & farm_type == "cabbage_farm" ~ 
        "Conventional\n(n=386)")) %>%
  mutate(
    value = round(value, 2),
    alt_short_variable_description = factor(short_variable_description, 
                                       levels = c("GCFI $1M or more", 
                                                  "GCFI $350,000 to $999,999",
                                                  "GCFI $75,000 to $349,999",
                                                  "GCFI $0 to $74,999")))

# Plot          
ggplot(gcfi_cat, aes(fill = short_variable_description, 
                      x = groupbyvar, 
                      y = value)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#B31B1B", "#9FAD9F", "#A2998B", "#073949")) + 
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(y = "Percent of operations") + 
  geom_text(aes(label = scales::percent(value)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "white") 
  
```

\newpage

## Percent of New York cabbage operations by scale (gross cash farm income)

```{r gcfi_bar_cabbage, fig.align="center", fig.width=7, fig.height=6, out.width="100%"}

## GCFI stacked bar chart 
# Format data 
gcfi_cat <- df %>% 
  select(!c(starts_with("se"), alt_short_variable_description)) %>%
  filter(str_detect(yvar, "gcfi_") & 
           farm_type=="cabbage_farm") %>%
  pivot_longer(
    cols = !c(farm_type, short_variable_description, yvar),
    names_to = "variable", 
    values_to = "value") %>% 
  mutate(
    groupbyvar = case_when(
      variable == "mean_organic" & farm_type == "leafy_green_farm" ~ 
        "Organic\n(n=172)",
       variable == "mean_conventional" & farm_type == "leafy_green_farm" ~ 
        "Conventional\n(n=609)", 
      variable == "mean_organic" & farm_type == "cabbage_farm" ~ 
        "Organic\n(n=111)",
       variable == "mean_conventional" & farm_type == "cabbage_farm" ~ 
        "Conventional\n(n=386)")) %>%
  mutate(
    value = round(value, 2),
    short_variable_description = factor(short_variable_description, 
                                       levels = c("GCFI $1M or more", 
                                                  "GCFI $350,000 to $999,999",
                                                  "GCFI $75,000 to $349,999",
                                                  "GCFI $0 to $74,999")))

# Plot          
ggplot(gcfi_cat, aes(fill = short_variable_description, 
                      x = groupbyvar, 
                      y = value)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#B31B1B", "#9FAD9F", "#A2998B", "#073949")) + 
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(y = "Percent of operations") + 
  geom_text(aes(label = scales::percent(value)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "white") 

rm(gcfi_cat)
```

# Variable production expenses

Shown as a total and as a percent of total variable expense.

-   Breeding stock, other livestock purchased or leased
-   Chemical expense
-   Customwork expense
-   Feed expense
-   Fertilizer expense
-   Gas, fuel, oil expense
-   Hired and contract labor expense
-   Repairs and maintenance expense
-   Other production expenses (Include storage and warehousing, marketing expenses, insurance, etc. Exclude health insurance premiums and payroll taxes)
-   Seeds and plants expense
-   Utilities expense
-   Total variable expense

```{r}

# List of variable expenses
varlist <- c("fert_chemical", "seedp", "animalpurchase", "feed", "fuelo", 
             "util", "mainr", "cwork", "labor", "totvarexp", 
             "fert_chemical_var_pct", "seedp_var_pct", 
             "animalpurchase_var_pct", "feed_var_pct", "fuelo_var_pct", 
             "util_var_pct", "mainr_var_pct", "cwork_var_pct", "labor_var_pct")

# Keep only variable expenses and variable expense as a percent of total variable expense
sum <- df %>% 
  filter(yvar %in% varlist)

# Sort from largest to smallest based on conventional
sum1 <- sum %>% 
  filter(
    !str_detect(yvar, "pct")) %>%
  arrange(mean_conventional)

sum2 <- sum %>% 
  filter(
    str_detect(yvar, "pct")) %>%
  arrange(mean_conventional)

sum <- bind_rows(sum1, sum2)

# Format data
sum <- sum %>% 
    mutate(across(c(mean_conventional:se_organic),
               ~case_when(
                  . > 1 ~ prettyNum(., big.mark = ",", digits = 2), 
                  . <= 1 ~ percent(., accuracy = 1.0))))  

# Put SE under mean in parenthesis
sum <- sum %>% 
  mutate(
    se_conventional = str_c("(", se_conventional, ")"), 
    se_organic = str_c("(", se_organic, ")"), 
    conventional = paste(mean_conventional, se_conventional), 
    organic = paste(mean_organic, se_organic)) %>% 
  select(farm_type, alt_short_variable_description, conventional, organic)

# Separate tables by commodity
sum_greens <- sum %>% 
  filter(farm_type=="leafy_green_farm") %>% 
  select(!farm_type) %>% 
  rename(`Conventional (n=609)` = conventional, 
         `Organic (n=172)` = organic, 
         Variable = alt_short_variable_description) 

sum_cabbage <- sum %>%
  filter(farm_type=="cabbage_farm") %>% 
  select(!farm_type) %>%
  rename(`Conventional (n=386)` = conventional, 
         `Organic (n=111)` = organic, 
         Variable = alt_short_variable_description)

```

\newpage

## Variable expenses - Leafy Green operations

```{r exp_greens}

kable(sum_greens) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Expenses are in 2022 dollars")

rm(sum_greens)
```

## Variable expenses - Cabbage operations

```{r exp_cabbage}

kable(sum_cabbage) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Expenses are in 2022 dollars")

rm(sum_cabbage)
```

## Variable expenses as a percent of total variable expense - Leafy Green operations

```{r expense_greens_stacked_bar, fig.align="center", fig.width=7, fig.height=6, out.width="100%"}
rm(sum)

# Create a stacked bar chart of variable expenses as a percent of total variable expense for conventional and organic operations, separate charts for leafy greens and cabbage operations 

# Keep variables and columns of interest
varlist <- c("fert_chemical_var_pct", "seedp_var_pct", 
             "animalpurchase_var_pct", "feed_var_pct", "fuelo_var_pct", 
             "util_var_pct", "mainr_var_pct", "cwork_var_pct", "labor_var_pct")

fig_df <- df %>% 
  select(!c(starts_with("se"), alt_short_variable_description)) %>%
  filter(yvar %in% varlist & 
           farm_type=="leafy_green_farm")

# Format data 
fig_df <- fig_df %>%
  pivot_longer(
    cols = !c(farm_type, short_variable_description, yvar),
    names_to = "variable", 
    values_to = "value") %>% 
  mutate(
    groupbyvar = case_when(
      variable == "mean_organic" & farm_type == "leafy_green_farm" ~ 
        "Organic\n(n=172)",
       variable == "mean_conventional" & farm_type == "leafy_green_farm" ~ 
        "Conventional\n(n=609)", 
      variable == "mean_organic" & farm_type == "cabbage_farm" ~ 
        "Organic\n(n=111)",
       variable == "mean_conventional" & farm_type == "cabbage_farm" ~ 
        "Conventional\n(n=386)")) 
# Plot
ggplot(fig_df, aes(fill = short_variable_description, 
                      x = groupbyvar, 
                      y = value)) + 
  geom_bar(position = "stack", stat = "identity")  +
  scale_fill_manual(values = c("#B31B1B", "#9FAD9F", "#A2998B", "#073949",
                               "#D47500", "#EF4035", "#F8981D", "#6EB43F", 
                               "#006699", "#222222", "#4B7B2B")) + 
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(y = "Percent of total variable expense") +
  geom_text(aes(label = label_percent(accuracy = 1)(value)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "white") 
```

## Variable expenses as a percent of total variable expense - Cabbage operations

```{r expense_cabbage_stacked_bar, fig.align="center", fig.width=7, fig.height=6, out.width="100%"}

# Create a stacked bar chart of variable expenses as a percent of total variable expense for conventional and organic operations, separate charts for leafy greens and cabbage operations 

# Keep variables and columns of interest
fig_df <- df %>% 
  select(!c(starts_with("se"), alt_short_variable_description)) %>%
  filter(yvar %in% varlist & 
           farm_type=="cabbage_farm")

# Format data 
fig_df <- fig_df %>%
  pivot_longer(
    cols = !c(farm_type, short_variable_description, yvar),
    names_to = "variable", 
    values_to = "value") %>% 
  mutate(
    groupbyvar = case_when(
      variable == "mean_organic" & farm_type == "leafy_green_farm" ~ 
        "Organic\n(n=172)",
       variable == "mean_conventional" & farm_type == "leafy_green_farm" ~ 
        "Conventional\n(n=609)", 
      variable == "mean_organic" & farm_type == "cabbage_farm" ~ 
        "Organic\n(n=111)",
       variable == "mean_conventional" & farm_type == "cabbage_farm" ~ 
        "Conventional\n(n=386)")) 

# Plot
ggplot(fig_df, aes(fill = short_variable_description, 
                      x = groupbyvar, 
                      y = value)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#B31B1B", "#9FAD9F", "#A2998B", "#073949",
                               "#D47500", "#EF4035", "#F8981D", "#6EB43F", 
                               "#006699", "#222222", "#4B7B2B")) + 
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(y = "Percent of total variable expense") +
  geom_text(aes(label = label_percent(accuracy = 1)(value)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "white") 
```

# Fixed production expenses

Shown as a total and as a percent of total fixed expense.

-   Interest paid on debts secured by real estate
-   Interest paid on debts not secured by real estate
-   Property taxes
-   Medical supplies, veterinary, custom services for livestock

Also include (but not as part of the total fixed expenses) - Depreciation expense claimed by the operation for all capital assets

```{r}

# List of variable expenses
varlist <- c("rent_land_buildings", "rent_lease_machinery", "property_tax", 
             "interest", "veterinary", "otherexp", "totfixedexp", 
             "rent_land_buildings_fixed_pct", "rent_lease_machinery_fixed_pct", 
             "property_tax_fixed_pct", "interest_fixed_pct", 
             "veterinary_fixed_pct", "otherexp_fixed_pct")

# Keep only variable expenses and variable expense as a percent of total variable expense
sum <- df %>% 
  filter(yvar %in% varlist)

# Sort from largest to smallest based on conventional
sum1 <- sum %>% 
  filter(
    !str_detect(yvar, "pct")) %>%
  arrange(mean_conventional)

sum2 <- sum %>% 
  filter(
    str_detect(yvar, "pct")) %>%
  arrange(mean_conventional)

sum <- bind_rows(sum1, sum2)

# Format data
sum <- sum %>% 
    mutate(across(c(mean_conventional:se_organic),
               ~case_when(
                  . > 1 ~ prettyNum(., big.mark = ",", digits = 2), 
                  . <= 1 ~ percent(., accuracy = 1.0))))  

# Put SE under mean in parenthesis
sum <- sum %>% 
  mutate(
    se_conventional = str_c("(", se_conventional, ")"), 
    se_organic = str_c("(", se_organic, ")"), 
    conventional = paste(mean_conventional, se_conventional), 
    organic = paste(mean_organic, se_organic)) %>% 
  select(farm_type, short_variable_description, conventional, organic)

# Separate tables by commodity
sum_greens <- sum %>% 
  filter(farm_type=="leafy_green_farm") %>% 
  select(!farm_type) %>% 
  rename(`Conventional (n=609)` = conventional, 
         `Organic (n=172)` = organic, 
         Variable = short_variable_description) 

sum_cabbage <- sum %>%
  filter(farm_type=="cabbage_farm") %>% 
  select(!farm_type) %>%
  rename(`Conventional (n=386)` = conventional, 
         `Organic (n=111)` = organic, 
         Variable = short_variable_description)

```

\newpage

## Variable expenses - Leafy Green operations

```{r fixedexp_greens}

kable(sum_greens) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Expenses are in 2022 dollars. Other expenses include storage and warehousing, marketing expenses, insurance, vehicle registration fees, etc. Excludes health insurance premiums and payroll taxes")

rm(sum_greens)
```

## Variable expenses - Cabbage operations

```{r fixedexp_cabbage}

kable(sum_cabbage) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Expenses are in 2022 dollars. Other expenses include storage and warehousing, marketing expenses, insurance, vehicle registration fees, etc. Excludes health insurance premiums and payroll taxes")

rm(sum_cabbage)
```

## Fixed expenses as a percent of total fixed expense - Leafy Green operations

```{r fixedexpense_greens_stacked_bar, fig.align="center", fig.width=7, fig.height=6, out.width="100%"}
rm(sum)

# Create a stacked bar chart of variable expenses as a percent of total variable expense for conventional and organic operations, separate charts for leafy greens and cabbage operations 

# Keep variables and columns of interest
varlist <- c("rent_land_buildings_fixed_pct", "rent_lease_machinery_fixed_pct", 
             "property_tax_fixed_pct", "interest_fixed_pct", 
             "veterinary_fixed_pct", "otherexp_fixed_pct")
fig_df <- df %>% 
  select(!starts_with("se")) %>%
  filter(yvar %in% varlist & 
           farm_type=="leafy_green_farm")

# Format data 
fig_df <- fig_df %>%
  pivot_longer(
    cols = !c(farm_type, short_variable_description, yvar),
    names_to = "variable", 
    values_to = "value") %>% 
  mutate(
    groupbyvar = case_when(
      variable == "mean_organic" & farm_type == "leafy_green_farm" ~ 
        "Organic\n(n=172)",
       variable == "mean_conventional" & farm_type == "leafy_green_farm" ~ 
        "Conventional\n(n=609)", 
      variable == "mean_organic" & farm_type == "cabbage_farm" ~ 
        "Organic\n(n=111)",
       variable == "mean_conventional" & farm_type == "cabbage_farm" ~ 
        "Conventional\n(n=386)")) 

# Plot
ggplot(fig_df, aes(fill = short_variable_description, 
                      x = groupbyvar, 
                      y = value)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#B31B1B", "#9FAD9F", "#A2998B", "#073949",
                               "#D47500", "#EF4035", "#F8981D", "#6EB43F", 
                               "#006699", "#222222", "#4B7B2B")) + 
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(y = "Percent of total variable expense") +
  geom_text(aes(label = label_percent(accuracy = 1)(value)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "white") 
```

## Fixed expenses as a percent of total fixed expense - Cabbage operations

```{r fixedexpense_cabbage_stacked_bar, fig.align="center", fig.width=7, fig.height=6, out.width="100%"}

# Create a stacked bar chart of variable expenses as a percent of total variable expense for conventional and organic operations, separate charts for leafy greens and cabbage operations 

# Keep variables and columns of interest
fig_df <- df %>% 
  select(!starts_with("se")) %>%
  filter(yvar %in% varlist & 
           farm_type=="cabbage_farm")

# Format data 
fig_df <- fig_df %>%
  pivot_longer(
    cols = !c(farm_type, short_variable_description, yvar),
    names_to = "variable", 
    values_to = "value") %>% 
  mutate(
    groupbyvar = case_when(
      variable == "mean_organic" & farm_type == "leafy_green_farm" ~ 
        "Organic\n(n=172)",
       variable == "mean_conventional" & farm_type == "leafy_green_farm" ~ 
        "Conventional\n(n=609)", 
      variable == "mean_organic" & farm_type == "cabbage_farm" ~ 
        "Organic\n(n=111)",
       variable == "mean_conventional" & farm_type == "cabbage_farm" ~ 
        "Conventional\n(n=386)")) 

# Plot
ggplot(fig_df, aes(fill = short_variable_description, 
                      x = groupbyvar, 
                      y = value)) + 
  geom_bar(position = "stack", stat = "identity") + 
  scale_fill_manual(values = c("#B31B1B", "#9FAD9F", "#A2998B", "#073949",
                               "#D47500", "#EF4035", "#F8981D", "#6EB43F", 
                               "#006699", "#222222", "#4B7B2B")) + 
  theme(legend.title = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(y = "Percent of total variable expense") +
  geom_text(aes(label = label_percent(accuracy = 1)(value)), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, color = "white") 
```

# Profitability

-   Net farm income
-   Operating profit margin
-   Return on assets

```{r}

# List of variable expenses
varlist <- c("roa", "opm", "infi")

# Keep only variable expenses and variable expense as a percent of total variable expense
sum <- df %>% 
  filter(yvar %in% varlist)

# Format data
sum <- sum %>% 
    mutate(across(c(mean_conventional:se_organic),
               ~case_when(
                  . > 1 ~ prettyNum(., big.mark = ",", digits = 2), 
                  . <= 1 ~ percent(., accuracy = 1.0))))  

# Put SE under mean in parenthesis
sum <- sum %>% 
  mutate(
    se_conventional = str_c("(", se_conventional, ")"), 
    se_organic = str_c("(", se_organic, ")"), 
    conventional = paste(mean_conventional, se_conventional), 
    organic = paste(mean_organic, se_organic)) %>% 
  select(farm_type, short_variable_description, conventional, organic)

# Separate tables by commodity
sum_greens <- sum %>% 
  filter(farm_type=="leafy_green_farm") %>% 
  select(!farm_type) %>% 
  rename(`Conventional (n=609)` = conventional, 
         `Organic (n=172)` = organic, 
         Variable = short_variable_description) 

sum_cabbage <- sum %>%
  filter(farm_type=="cabbage_farm") %>% 
  select(!farm_type) %>%
  rename(`Conventional (n=386)` = conventional, 
         `Organic (n=111)` = organic, 
         Variable = short_variable_description)

```

## Profitability - Leafy Green operations

```{r profitability_greens}

kable(sum_greens) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Net farm income is in 2022 dollars")

```

## Profitability - Cabbage operations

```{r profitability_cabbage}

kable(sum_cabbage) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Net farm income is in 2022 dollars")

```

# Land use practices

-   Cropland acres planted to a cover crop
-   No till acres
-   Reduced till acres

```{r}

# List of variable expenses
varlist <- c("cover_crop", "reduced_till", "no_till")

# Keep only variable expenses and variable expense as a percent of total variable expense
sum <- df %>% 
  filter(yvar %in% varlist)

# Format data
sum <- sum %>% 
    mutate(across(c(mean_conventional:se_organic),
               ~case_when(
                  . > 1 ~ prettyNum(., big.mark = ",", digits = 2), 
                  . <= 1 ~ percent(., accuracy = 1.0))))  

# Put SE under mean in parenthesis
sum <- sum %>% 
  mutate(
    se_conventional = str_c("(", se_conventional, ")"), 
    se_organic = str_c("(", se_organic, ")"), 
    conventional = paste(mean_conventional, se_conventional), 
    organic = paste(mean_organic, se_organic)) %>% 
  select(farm_type, short_variable_description, conventional, organic)

# Separate tables by commodity
sum_greens <- sum %>% 
  filter(farm_type=="leafy_green_farm") %>% 
  select(!farm_type) %>% 
  rename(`Conventional (n=609)` = conventional, 
         `Organic (n=172)` = organic, 
         Variable = short_variable_description) 

sum_cabbage <- sum %>%
  filter(farm_type=="cabbage_farm") %>% 
  select(!farm_type) %>%
  rename(`Conventional (n=386)` = conventional, 
         `Organic (n=111)` = organic, 
         Variable = short_variable_description)

```

## Land Use Practices - Leafy Green operations

```{r landuse_greens}

kable(sum_greens) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Net farm income is in 2022 dollars")

```

## Land Use Practices - Cabbage operations

```{r landuse_cabbage}

kable(sum_cabbage) %>% 
  kable_styling() %>%
  footnote(general = "Mean (SE). Net farm income is in 2022 dollars")

```
