---
format: html
editor: visual
execute: 
  message: false
  warning: false
  eval: true
---

```{r}

library(tidyverse)
library(openxlsx)
library(readxl)
library(janitor)

```

[Final data: Pricing for beef along the supply chain .](https://github.com/FoodSystemsModeling/CityFoodPolicy/tree/main/beef/data_final/beef_price.xlsx)

### Differentiated

ADD

```{r}

diff_price <- tibble(
  supply_chain_location = c(
    "Choice beef gross farm value", 
    "Choice beef retail value",
    "Choice beef wholesale value"
  ),
  beef_price_kg = "need data", 
  organic_beef_price_kg = "need data"
)

```

### Commodity

We use [Meat Price Spreads data from the USDA Economic Research Service, Historical monthly price spread data for beef, pork, broilers](https://www.ers.usda.gov/data-products/meat-price-spreads).

Here we focus on price data for the last 10 years, where available. All prices are reported in 2024 dollars.

#### Monthly prices along suppy chain

```{r}

# Import USDA ERS price spread data
df <- read_csv("../data_raw/history.csv") %>% 
  clean_names()

# Keep beef data since 2014
df <- df %>% 
  filter(year>=2014 & 
           str_detect(data_item, "beef"))


# Keep data of interest
df <- df %>% 
  filter(data_item %in% c(
    "Choice beef gross farm value", 
    "Choice beef wholesale value",
    "Choice beef retail value"
  ))

# Make year-month date
df <- df %>% 
  mutate(
    date = make_date(year, month_number, 1)
  )

```

##### PPI

For prices at the farmgate, we use the [producer price index by commodity: Farm products: Slaughter Cattle](https://fred.stlouisfed.org/series/WPU01830131)("WPU0131.csv") to put all beef price data in 2024 dollars.

Citation: U.S. Bureau of Labor Statistics, Producer Price Index by Commodity: Farm Products: Slaughter Cattle \[WPU0131\], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/WPU0131, August 17, 2026.

For retaila and wholesale prices, we use the [Producer Price Index by Commodity: Processed Foods and Feeds: Meats (WPU0221)](https://fred.stlouisfed.org/series/WPU0221)("WPU0221.csv").

Citation: U.S. Bureau of Labor Statistics, Producer Price Index by Commodity: Processed Foods and Feeds: Meats \[WPU0221\], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/WPU0221, August 17, 2026.

We compute an average yearly PPI and then use that to convert all dollars into 2024 dollars. We keep data starting in 2014.

```{r}

# Import producer ppi
producer_ppi <- read_csv("../data_raw/WPU0131.csv") %>% 
  filter(observation_date>= as.Date("2014-01-01"))

# PPI get average per year
producer_ppi <- producer_ppi %>% 
  mutate(year = year(observation_date)) %>% 
  group_by(year) %>% 
  summarise(ppi = mean(WPU0131, na.rm = TRUE)) %>% 
  mutate(producer_ppi_index_2024 = ppi[year == 2024] / ppi) %>% 
  select(-ppi)


# Import retail/wholesale ppi
retail_wholesale_ppi <- read_csv("../data_raw/WPU0221.csv") %>% 
  filter(observation_date>= as.Date("2014-01-01"))

# PPI get average per year
retail_wholesale_ppi <- retail_wholesale_ppi %>% 
  mutate(year = year(observation_date)) %>% 
  group_by(year) %>% 
  summarise(ppi = mean(WPU0221, na.rm = TRUE)) %>% 
  mutate(retail_wholesale_ppi_index_2024 = ppi[year == 2024] / ppi) %>% 
  select(-ppi)

# Join together
ppi <- full_join(producer_ppi, retail_wholesale_ppi)


# Join with data 
df <- df %>% 
  left_join(ppi)

rm(ppi, producer_ppi, retail_wholesale_ppi)

# Covert prices
df <- df %>% 
  mutate(
    real_beef_price_2024dollars = case_when(
      data_item=="Choice beef gross farm value" ~ value*producer_ppi_index_2024, 
      TRUE ~ value*retail_wholesale_ppi_index_2024
    )
  )

```

#### Figures

```{r}

# Choice beef gross farm value
ggplot(df %>% 
         filter(data_item=="Choice beef gross farm value"), 
       aes(x = date, 
           y = real_beef_price_2024dollars)) +
  geom_line() +
  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", color = "grey50") +
  labs(
    x = NULL,
    y = "Price (2024 Cents per pound of retail equivalent)",
    title = "Real monthly Choice beef gross farm value"
  ) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y",
               limits = as.Date(c("2014-01-01", "2024-12-31"))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# Choice beef wholesale value
ggplot(df %>% 
         filter(data_item=="Choice beef wholesale value"), 
       aes(x = date, 
           y = real_beef_price_2024dollars)) +
  geom_line() +
  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", color = "grey50") +
  labs(
    x = NULL,
    y = "Price (2024 Cents per pound of retail equivalent)",
    title = "Real monthly Choice beef wholesale value"
  ) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y",
               limits = as.Date(c("2014-01-01", "2024-12-31"))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# Choice beef retail value
ggplot(df %>% 
         filter(data_item=="Choice beef retail value"), 
       aes(x = date, 
           y = real_beef_price_2024dollars)) +
  geom_line() +
  geom_smooth(method = "loess", se = FALSE, linetype = "dashed", color = "grey50") +
  labs(
    x = NULL,
    y = "Price (2024 Cents per pound of retail equivalent)",
    title = "Real monthly Choice beef retail value"
  ) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y",
               limits = as.Date(c("2014-01-01", "2024-12-31"))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Organic

Using data from the [USDA AMS Livestock, Poultry, and Grain Market News, New York Dept of Ag Mrkt News, Feb 5, 2025](https://www.ams.usda.gov/mnreports/ams_3789.pdf), we estimate the increased cost of organic beef at the farmgate.

Data are pulled from Slaughter Cattle, Dairy Cows, we use DAIRY COWS - Lean 85-90% (Per Cwt / Actual Wt) as it has the largest sample size. The average price per cwt/actual wt for conventional is 116.98 and for organic is 118.78. Organic is 1.5% higher than conventional. We apply a 1.8% premium on farmgate organic prices.

Retail prices are found from the [USDA AMS Livestock, Poultry, & Grain Market News, Weekly Grocery Store Beef Feature Activity, August 14, 2026](https://www.ams.usda.gov/mnreports/ams_3228.pdf). We apply the prices differences of conventional to organic to both retail and wholesale. We use the price of ground beef, Ground Beef 80-89%, 1-2 Lbs from the Previous Week (PW).

Conventional: 6.26 USDA Organic, fresh: 6.99

We apply a 11.7% premium on all retail and wholesale organic prices.

```{r}

# Define how much to inflate prices for organic
organic_farmgate <- (118.78-116.98)/116.98
organic_retail_wholesale <- (6.99-6.26)/6.26

# Define organic price
df <- df %>% 
  mutate(
    organic_real_beef_price_2024dollars =
      case_when(
        data_item == "Choice beef gross farm value" ~ real_beef_price_2024dollars + (real_beef_price_2024dollars*organic_farmgate), 
        TRUE ~ real_beef_price_2024dollars + (real_beef_price_2024dollars*organic_retail_wholesale)
  )
)
```

### Final table

For the model, [we include one worksheet with differentiated prices, one with commodity beef prices at each level of the supply chain](https://github.com/FoodSystemsModeling/CityFoodPolicy/tree/main/beef/data_final/beef_price.xlsx). These prices are an average of the prices from 2022-2024. All prices are in 2024 dollars.

Convert all prices into \$/kg. 1 kg = 2.20462 lb

cents/kg=cents/lb×2.20462

\$/kg=100cents/lb×2.20462

```{r}

# Commodity
sum <- df %>% 
  filter(year %in% c(2022, 2023, 2024)) %>% 
  group_by(data_item) %>%
  summarise(
    beef_price_cents_lb_2024dollars = mean(real_beef_price_2024dollars), 
    organic_beef_price_cents_lb_2024dollars = 
      mean(organic_real_beef_price_2024dollars)) %>%
  mutate(
    beef_price_kg = (beef_price_cents_lb_2024dollars*2.20462) / 100, 
    organic_beef_price_kg = (organic_beef_price_cents_lb_2024dollars*2.20462) / 100
    ) 
    
# Keep data of interest
sum <- sum %>% 
  rename(supply_chain_location = data_item) %>% 
  select(supply_chain_location, 
         beef_price_kg, organic_beef_price_kg)
    

# Create workbook
wb <- createWorkbook()

# Add differentiated
addWorksheet(wb, "differentiated_price")
writeData(wb, "differentiated_price", diff_price)

## Add commodity
addWorksheet(wb, "commodity_price")
writeData(wb, "commodity_price", sum)

# Add info
info <- tibble(
  sheet_name = c(
    "differentiated_price", "commodity_price"), 
  sheet_description = c(
    "prices along the supply chain for differentiated culled beef", 
    "average price per kg from 2022-2024 (in 2024 dollars) for Choice beef gross farm value, Choice beef wholesale value, and Choice beef retail value"), 
  source = c(
    "??", 
    "Meat Price Spreads data from the USDA Economic Research Service, Historical monthly price spread data for beef, pork, broilers (https://www.ers.usda.gov/data-products/meat-price-spreads), we inflate organic prices using USDA AMS Livestock, Poultry, and Grain Market News, New York Dept of Ag Mrkt News, Feb 5, 2025 (https://www.ams.usda.gov/mnreports/ams_3789.pdf) for farmgate prices and Retail prices are found from the USDA AMS Livestock, Poultry, & Grain Market News, Weekly Grocery Store Beef Feature Activity, August 14, 2026(https://www.ams.usda.gov/mnreports/ams_3228.pdf) for wholesale and retail prices")
)

# Add worksheet
addWorksheet(wb, "info")
writeData(wb, "info", info)

# Save workbook 
saveWorkbook(wb, "../data_final/beef_price.xlsx", 
             overwrite = TRUE)

```
